---
title: "bachelor thesis"
author: "0juliat0"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r, include=FALSE}
library(plotly)
library(ggplot2)
library(CellChat)
library(dplyr)
library(tidyr)
library(Matrix)
library(ggplot2)
library(pheatmap)
library(Seurat)
```

Load data
```{r}
file_path_t <- "~/rocker/RDEB/ccc_cellchat.t.2.rds"
cellchat.t <- readRDS(file_path_t)
file_path_s <- "~/rocker/RDEB/ccc_cellchat.s.2.rds"
cellchat.s <- readRDS(file_path_s)
```

Circle diagramm for skin, SCC
```{r}
cellchat.t@netP$pathways
pathways.show.t <- c("COLLAGEN", "LAMININ", "MIF", "MHC-I", "FN1", "CD99", "THBS", "VISFATIN", "MHC-II", "APP", "CLEC", "MK", "ADGRE5", "GALECTIN", "CXCL", "TGFb", "PARs", "SEMA4", "PTPRM", "SEMA3", "TENASCIN", "CD45", "ANNEXIN", "PTN", "ICAM", "ITGB2", "GAS", "EGF", "PECAM1", "ANGPTL", "TNF", "CD22", "JAM", "CD46", "VEGF", "NOTCH", "THY1", "PERIOSTIN", "CCL", "SELE", "MPZ", "ncWNT", "NEGR", "ALCAM", "CD6", "ANGPT", "NECTIN", "COMPLEMENT", "VCAM", "SELL", "FGF", "PDGF", "PROS", "CADM", "HSPG", "CALCR", "CD226", "BAFF", "CD86", "NCAM", "IL16", "IL1", "SEMA6", "EPHB", "BMP", "EPHA", "LCK", "TIGIT", "IFN-II", "GRN", "ESAM", "NRXN", "KIT", "CDH5", "CSF", "SELPLG", "CDH", "EDN", "BTLA", "IL2", "CNTN", "CD70", "TWEAK", "IGF", "APRIL", "TRAIL", "VISTA", "NRG", "APJ", "RELN", "IL6", "CD34", "HGF", "L1CAM", "SN", "CD137", "WNT")

cellchat.s@netP$pathways
pathways.show.s <- c("COLLAGEN", "LAMININ", "MHC-I", "MIF", "FN1", "CD99", "THBS", "VISFATIN", "CLEC", "APP", "ADGRE5", "MHC-II", "CXCL", "ICAM", "SELE", "MK", "CD45", "PTN", "PTPRM", "ITGB2", "PARs", "CD22", "ALCAM", "CD6", "TGFb", "ANGPTL", "VEGF", "JAM", "CD46", "ANNEXIN", "PECAM1", "GAS", "FGF", "SEMA3", "MPZ", "CADM", "NOTCH", "TENASCIN", "VCAM", "NEGR", "PERIOSTIN", "NECTIN", "IL6", "ANGPT", "SEMA4", "THY1", "EGF", "TIGIT", "TNF", "PDGF", "CCL", "SEMA7", "CALCR", "NCAM", "IL2", "ncWNT", "IFN-II", "EPHA", "PROS", "IL16", "LCK", "EPHB", "CD226", "HSPG", "SELL", "KIT", "BMP", "ESAM", "CD70", "CDH5", "LIGHT", "CD86", "DESMOSOME", "BAFF", "SEMA6", "TRAIL", "COMPLEMENT", "IGF", "CDH1", "SELPLG", "IL1", "FASLG", "EDN", "TWEAK", "NRXN", "IL10", "CD34", "CSF3", "LIFR", "CSF", "NRG", "L1CAM", "CDH", "WNT", "VEGI")

# Output PDF setup
pdf("~/rocker/RDEB/thesis/Chord_diagram_SCC.pdf")
netVisual_aggregate(cellchat.t, signaling = pathways.show.t, layout = "chord")

# Output PDF setup
pdf("~/rocker/RDEB/thesis/Chord_diagram_Skin.pdf")
netVisual_aggregate(cellchat.s, signaling = pathways.show.s, layout = "chord")
```
  
Heatmap SCC, heatmap skin
```{r}
# Set up PDF output for SCC, saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_SCC_Interaction_count.pdf", width = 16, height = 8)

# Interaction count heatmap for SCC
gg1_tumor <- netVisual_heatmap(cellchat.t, measure = "count")
plot(gg1_tumor)

# Set up PDF output for Skin, also saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Skin_Interaction_count.pdf", width = 16, height = 8)

# Interaction count heatmap for Skin
gg1_skin <- netVisual_heatmap(cellchat.s, measure = "count")

plot(gg1_skin)

```

There are no T naive CD4 in skin so lets exclude them from analysis 
```{r}
# Define the identifier for naive CD4 T cells
naive_CD4_identifier <- "T naive CD4"

# Filter out interactions involving naive CD4 T cells from the SCC dataset
cellchat.t.sC <- subsetCommunication(cellchat.t, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | 
                            target == naive_CD4_identifier))

# Filter out interactions involving naive CD4 T cells from the skin dataset
cellchat.s.sC <- subsetCommunication(cellchat.s, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | 
                            target == naive_CD4_identifier))

# merge the two data frames
merged_data <- merge(cellchat.t.sC, cellchat.s.sC, 
                     by = c("source", "target", "ligand", "receptor", 
                            "interaction_name","interaction_name_2", 
                            "pathway_name", "annotation", "evidence"),
                     all = TRUE)

```
Replace NA values with 0
```{r}
merged_data <- merged_data %>%
  mutate(
    prob.t = ifelse(is.na(prob.x), 0, prob.x),
    prob.s = ifelse(is.na(prob.y), 0, prob.y),
    pval.t = ifelse(is.na(pval.x), 0, pval.x),
    pval.s = ifelse(is.na(pval.y), 0, pval.y)
  ) %>%
  select(-prob.x, -prob.y, -pval.x, -pval.y) # delete columns 
```

Calculating the difference between probabilities SCC - skin
```{r}
merged_data <- merged_data %>%
  mutate(prob.delta = prob.t - prob.s) %>%
  arrange(desc(abs(prob.delta)))
```

Count interactions for skin and SCC (actual numeric difference )
```{r}
# Create the summary data frame with the desired columns
merged_df.interactions <- merged_data %>%
  group_by(source, target) %>%
  summarise(
    Tumor_Interactions = sum(prob.t > 0),
    Skin_Interactions = sum(prob.s > 0),
    Interaction_Difference = Tumor_Interactions - Skin_Interactions,
    Interaction_Difference_abs = sum(xor(prob.t == 0, prob.s == 0)),
    .groups = 'drop'
  )

# Print the interaction summary data frame
print(merged_df.interactions)
```

Bubble plot based on counts difference
```{r}
# Create a point plot with color and size adjustments
point_plot <- ggplot(merged_df.interactions, aes(x = target, y = source)) +
  geom_point(aes(size = Interaction_Difference_abs, color = Interaction_Difference), shape = 20) +
  scale_size_continuous(name = "Absolute Interaction Difference", range = c(1, 11)) + 
  scale_color_gradient2(name = "Interactions Difference",
                        high = rgb(178/255, 24/255, 43/255), 
                        mid = "white",
                        low = rgb(33/255, 102/255, 172/255),
                        midpoint = 0) + # Diverging color scale
  labs(title = "Difference in Number of Interactions (SCC vs Skin)",
       x = "Target cells (receptor)",
       y = "Source cells (ligand)",
       color = "Interactions Difference",
       size = "Abs Interaction Diff") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

# Print the point plot
print(point_plot)

# Save the plot
ggsave("~/rocker/RDEB/thesis/bubble_plot_based_on_interactions_difference.png",
       plot = point_plot,
       width = 10,
       height = 8,
       dpi = 300)
```

Number of interactions per source cell (ligand) in SCC - skin
```{r}
# Compute interaction counts for each dataset, excluding T naive CD4
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>%
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% # no need to exclude, since in skin there is nothing 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'source' (27 levels - celltypes) column
combined_data.diff <- full_join(df.t, df.s, by = "source")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each source: skin - SCC
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions - combined_data.diff$Skin_Interactions

# Display the table (screenshot saved)
print(combined_data.diff, n = 27)# Display the table (screenshot saved)
```

Scatter plot source
```{r}
scatter_plot <- ggplot(
  combined_data.diff, 
  aes(x = Tumor_Interactions, 
      y = Skin_Interactions, 
      label = source)) +
  geom_point(
    aes(color = source), size = 3
    ) +
  geom_text(
    aes(label = source), size = 3
    ) +
  geom_smooth(
    method = "lm",
    se = TRUE
    ) +
  labs(
    title = "Number of Interactions per source Cell", 
    x = "SCC Interactions", 
    y = "Skin Interactions"
    ) +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)

ggsave("~/rocker/RDEB/thesis/scatter_plot_source.png", 
       plot = scatter_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Number of interactions per target cell (receptor) in SCC - skin
```{r}
# Compute interaction counts for each dataset
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'target' (27 levels - celltypes) column
combined_data.diff <- full_join(
  df.t,
  df.s, 
  by = "target")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each target: SCC - skin
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions
print (combined_data.diff, n = 27) # Display the table (screenshot saved)
```

Scatter plot target
```{r}
scatter_plot <- ggplot(combined_data.diff, 
                       aes(
                         x = Tumor_Interactions, 
                         y = Skin_Interactions, 
                         label = target)
                       ) + 
  geom_point(aes(color = target), size = 3) +
  geom_text(aes(label = target), size = 3) +
  geom_smooth(method = "lm",se = TRUE) +
  labs(
    title = "Number of Interactions per target Cell", 
       x = "SCC Interactions", 
       y = "Skin Interactions"
    ) +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)
ggsave("~/rocker/RDEB/thesis/scatter_plot_target.png", 
       plot = scatter_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Plotting top 10 interactions with biggest differences in SCC - skin
```{r}
# Combining data SCC and skin 
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>%
  dplyr::summarise(Tumor_Interactions = n())

df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>% 
  dplyr::summarise(Skin_Interactions = n())

# Merge datasets by the 'source' and 'target'
combined_data.diff <- full_join(
  df.t, 
  df.s, 
  by = c("source", "target")
  )

# Replace NAs with 0
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each source-target pair: skin - SCC
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions

# Filter to get the top 10 interactions by absolute difference
top_10_diff <- combined_data.diff %>%
  dplyr::ungroup() %>% # Calculate differences without grouping by 'source'
  dplyr::arrange(-abs(Interaction_Difference)) %>%
  dplyr::slice_head(n = 10)

print(top_10_diff)

# Plot top 10 interactions with biggest differences 
difference_plot <- ggplot(
  top_10_diff, 
  aes(
    x = paste(source, target, sep = "-"), 
    y = Interaction_Difference,
    fill = Interaction_Difference > 0)
  ) +
  geom_bar(
    stat = "identity",
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255),
      "FALSE" = rgb(33/255, 102/255, 172/255)
      ),
                    labels = c(
                      "TRUE" = "SCC",
                      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(
    title = "Top 10 Differences in Interactions (SCC - Skin)", 
    y = "Difference in Interactions", 
    x = "Interaction (Source-target)") +
  theme(axis.text.x = element_text(
    angle = 90, 
    hjust = 1)
    )

print(difference_plot)

ggsave("~/rocker/RDEB/thesis/Top_10_Differences_in_Interactions.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Function to process interactions between sorce and target celltypes 
```{r}
plot_interaction <- function(
    df, 
    source_value, 
    target_value) {
  df_filtered <- df %>%
    filter(source == source_value & target == target_value) %>%
    slice_max(
      order_by = abs(prob.delta), 
      n = 40) %>%
    select(-c(pval.t, pval.s))

  # Create a label from the source and target values
  interaction_label <- paste(
    source_value, 
    target_value, 
    sep = " to ")

  # Pivot and plot using the new label
  df_filtered %>%
    pivot_longer(cols = c("prob.s", "prob.t"), 
                 names_to = "variable", 
                 values_to = "value") %>%
    ggplot(aes(
      x = interaction_name_2, 
      y = value, 
      fill = variable)
      ) +
    geom_col(position = "dodge") +
    scale_fill_manual(
      values = c(
      "prob.s" = rgb(33/255, 102/255, 172/255), 
      "prob.t" = rgb(178/255, 24/255, 43/255)
      )
      ) +
    labs(
      x = interaction_label, 
      y = "Probability Delta", 
      fill = "Probability"
      ) +
    theme(axis.text.x = element_text(
      angle = 90, 
      hjust = 1, 
      vjust = 0.5)
      )
}
```

Interaction based on cell types
```{r}
# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, 
           target == current_target)

  # Plot
  p <- ggplot(
    current_data, 
    aes(
      x = ligand, 
      y = receptor, 
      fill = prob.delta)
    ) +
    geom_tile() +
    scale_fill_gradient2(
      low = rgb(33/255, 102/255, 172/255), 
      high = rgb(178/255, 24/255, 43/255), 
      mid = "grey", 
      midpoint = 0,
      limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                 max(current_data$prob.delta, na.rm = TRUE)),
      name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for",
                       current_source,
                       "to",
                       current_target),
         x = "Ligand",
         y = "Receptor"
         ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
      angle = 90, 
      vjust = 0.5
      ),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("~/rocker/RDEB/thesis/interaction_plots/",
                      current_source,
                      "_to_",
                      current_target, 
                      ".pdf"
                      )
  ggsave(
    file_name,
    plot = p,
    width = 10,
    height = 10
    )
  
  # Extracting the relevant interactions from merged_data for the current interaction pair
  filtered_data <- current_data %>%
    distinct(pathway_name)  # Getting unique pathway names involved in these interactions

  # Print the pathways involved in the current interaction pair
  cat("Pathways for interaction", current_source, "to", current_target, ":\n")
  print(filtered_data)

  # Check if these pathways are in the provided pathways lists
  pathways_involved_t <- filtered_data$pathway_name %in% pathways.show.t
  pathways_involved_s <- filtered_data$pathway_name %in% pathways.show.s

  # Print pathways involved both in tumor and skin for the current interaction pair
  cat("Pathways in Tumor:\n")
  print(filtered_data$pathway_name[pathways_involved_t])
  cat("Pathways in Skin:\n")
  print(filtered_data$pathway_name[pathways_involved_s])
}

# List of interactions
interactions <- list(
  "Fibroblast_32 to Fibroblast_32" = list(
    "tumor" = c("COLLAGEN", "NEGR", "MK", "LAMININ", "TENASCIN", "SEMA4", "FN1", "VISFATIN", "FGF", "THBS", "ncWNT", "ANGPTL", "TGFb", "GAS", "CD99", "PTPRM", "JAM", "CXCL", "PERIOSTIN", "SEMA3", "PROS", "CADM", "PDGF", "CNTN", "PTN", "NECTIN", "TWEAK", "BMP", "MIF", "MPZ", "IGF", "WNT"),
    "skin" = c("COLLAGEN", "NEGR", "MK", "LAMININ", "TENASCIN", "SEMA4", "FN1", "VISFATIN", "FGF", "THBS", "ncWNT", "ANGPTL", "TGFb", "GAS", "CD99", "PTPRM", "JAM", "CXCL", "PERIOSTIN", "SEMA3", "PROS", "CADM", "PDGF", "PTN", "NECTIN", "TWEAK", "BMP", "MIF", "MPZ", "IGF", "WNT")
  ),
  "Fibroblast_36 to Fibroblast_36" = list(
    "tumor" = c("COLLAGEN", "ANGPTL", "LAMININ", "MK", "THBS", "MPZ", "VISFATIN", "JAM", "NEGR", "IL6", "SEMA3", "FN1", "SEMA4", "FGF", "GAS", "TENASCIN", "ADGRE5", "PROS", "CXCL", "NOTCH", "CD46", "NECTIN", "PTN", "TWEAK", "CD99", "PERIOSTIN", "HSPG", "PTPRM", "EPHA", "TGFb", "PDGF", "EPHB", "CADM", "MIF", "IGF", "WNT"),
    "skin" = c("COLLAGEN", "ANGPTL", "LAMININ", "MK", "THBS", "MPZ", "VISFATIN", "JAM", "NEGR", "IL6", "SEMA3", "FN1", "SEMA4", "FGF", "GAS", "TENASCIN", "ADGRE5", "PROS", "CXCL", "NOTCH", "CD46", "NECTIN", "PTN", "TWEAK", "CD99", "PERIOSTIN", "HSPG", "PTPRM", "EPHA", "TGFb", "PDGF", "EPHB", "CADM", "MIF", "IGF", "WNT")
  ),
  "Fibroblast_32 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "MK", "VISFATIN", "MIF", "FN1", "CXCL", "SEMA4", "CALCR", "ANGPTL", "CD46", "CCL", "PTN", "ncWNT", "VEGF", "PTPRM", "CD99", "SEMA3", "NECTIN", "APJ", "PDGF", "APP", "IGF"),
    "skin" = c("COLLAGEN", "LAMININ", "MK", "VISFATIN", "MIF", "FN1", "CXCL", "SEMA4", "CALCR", "ANGPTL", "CD46", "CCL", "PTN", "ncWNT", "VEGF", "PTPRM", "CD99", "SEMA3", "NECTIN", "PDGF", "APP", "IGF")
  ),
  "Endothel_vasc._31 to Fibroblast_32" = list(
    "tumor" = c("COLLAGEN", "SELE", "LAMININ", "VISFATIN", "JAM", "ANGPT", "THBS", "GALECTIN", "SEMA4", "CD99", "SEMA6", "NOTCH", "PTPRM", "FN1", "EDN", "TENASCIN", "TGFb", "PROS", "PERIOSTIN", "SEMA3", "VCAM", "CXCL", "BMP", "MIF", "GAS", "MPZ", "EGF", "PDGF", "WNT"),
    "skin" = c("COLLAGEN", "SELE", "LAMININ", "VISFATIN", "JAM", "ANGPT", "THBS", "SEMA4", "CD99", "SEMA6", "NOTCH", "PTPRM", "FN1", "EDN", "TENASCIN", "TGFb", "PROS", "PERIOSTIN", "SEMA3", "VCAM", "CXCL", "BMP", "MIF", "GAS", "MPZ", "EGF", "PDGF", "WNT") 
  ),
  "Fibroblast_32 to Monocyte" = list(
    "tumor" = c("FN1", "COLLAGEN", "CD99", "LAMININ", "COMPLEMENT", "TENASCIN", "MK", "THY1", "MIF", "THBS", "CXCL", "ANNEXIN", "IL6", "FGF", "ANGPTL", "GAS", "CSF", "MHC-I", "SEMA3", "SEMA4", "MHC-II", "JAM", "MPZ", "IL16", "ICAM", "PROS", "NECTIN", "PTN", "APP", "TGFb", "VISFATIN", "VEGF"),
    "skin" = c("FN1", "COLLAGEN", "CD99", "LAMININ", "COMPLEMENT", "TENASCIN", "MK", "THY1", "MIF", "THBS", "CXCL", "ANNEXIN", "IL6", "FGF", "ANGPTL", "GAS", "CSF", "MHC-I", "SEMA3", "SEMA4", "MHC-II", "JAM", "MPZ", "IL16", "ICAM", "PROS", "NECTIN", "PTN", "APP", "TGFb", "VISFATIN", "VEGF")
  ),
  "Fibroblast_20 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "CD99", "VISFATIN", "MIF", "FN1", "VEGF", "CD46", "SEMA4", "ADGRE5", "CCL", "ESAM", "PDGF", "ANGPT", "APP", "PTPRM"),
    "skin" = c("COLLAGEN", "LAMININ", "CD99", "VISFATIN", "MIF", "FN1", "VEGF", "CD46", "SEMA4", "ADGRE5", "CCL", "ESAM", "PDGF", "ANGPT", "APP", "PTPRM")
  ),
  "Fibroblast_32 to Fibroblast_20" = list(
    "tumor" = c("LAMININ", "MK", "CD99", "CD46", "COLLAGEN", "GAS", "PERIOSTIN", "THBS", "VISFATIN", "ncWNT", "NEGR", "CALCR", "MPZ", "HSPG", "JAM", "FN1", "PTN", "PROS", "NECTIN", "PDGF", "CSF", "PTPRM", "BMP", "EPHA", "IGF"),
    "skin" = c("LAMININ", "MK", "CD99", "CD46", "COLLAGEN", "GAS", "PERIOSTIN", "THBS", "VISFATIN", "ncWNT", "NEGR", "CALCR", "MPZ", "HSPG", "JAM", "FN1", "PTN", "PROS", "NECTIN", "PDGF", "CSF", "PTPRM", "BMP", "EPHA", "IGF")
  ),
  "Fibroblast_18 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "VISFATIN", "FN1", "ncWNT", "SEMA4", "CXCL", "VEGF", "ANGPTL", "MIF", "APP", "CD46", "SEMA3", "PTN", "MK", "CCL", "NECTIN", "PTPRM", "APJ", "CALCR", "PDGF", "CD99", "IGF"),
    "skin" = c("COLLAGEN", "LAMININ", "VISFATIN", "FN1", "ncWNT", "SEMA4", "CXCL", "VEGF", "ANGPTL", "MIF", "APP", "CD46", "SEMA3", "PTN", "MK", "CCL", "NECTIN", "PTPRM", "CALCR", "PDGF", "CD99", "IGF")
  ),
  "Fibroblast_18 to Fibroblast_36" = list(
    "tumor" = c("COLLAGEN", "THBS", "LAMININ", "ANGPTL", "FGF", "ncWNT", "VISFATIN", "SEMA4", "IL6", "NEGR", "SEMA3", "GAS", "MPZ", "FN1", "CADM", "TENASCIN", "PTN", "TGFb", "JAM", "MK", "PROS", "PTPRM", "EPHA", "NECTIN", "CD46", "PDGF", "HSPG", "CD99", "MIF", "PERIOSTIN", "CXCL", "IGF"),
    "skin" = c("COLLAGEN", "THBS", "LAMININ", "ANGPTL", "FGF", "ncWNT", "VISFATIN", "SEMA4", "IL6", "NEGR", "SEMA3", "GAS", "MPZ", "FN1", "CADM", "TENASCIN", "PTN", "TGFb", "JAM", "MK", "PROS", "PTPRM", "EPHA", "NECTIN", "CD46", "PDGF", "HSPG", "CD99", "MIF", "PERIOSTIN", "CXCL", "IGF")
  ),
  "Endothel to Endothel" = list(
    "tumor" = c("COLLAGEN", "MIF", "LAMININ", "VISFATIN", "CD99", "SEMA4", "FN1", "PECAM1", "CXCL", "CD46", "CCL", "APP", "ESAM", "PTPRM", "ANGPT", "CDH5"),
    "skin" = c("COLLAGEN", "MIF", "LAMININ", "VISFATIN", "CD99", "SEMA4", "FN1", "PECAM1", "CXCL", "CD46", "CCL", "APP", "ESAM", "PTPRM", "ANGPT", "CDH5")
  )
)

# Find common pathways for each interaction pair
common_pathways_each_pair <- lapply(interactions, function(x) {
  intersect(x$tumor, x$skin)
})

# Find common pathways across all interaction pairs
common_pathways_across_all <- Reduce(intersect, common_pathways_each_pair)

# Print common pathways across all interaction pairs
cat("Common pathways across all interaction pairs:\n")
print(common_pathways_across_all)

# Find pathways that are present in either tumor OR skin for each interaction pair
all_pathways_each_pair <- lapply(interactions, function(x) {
  union(x$tumor, x$skin)
})

# Print pathways for each interaction pair
for (interaction in names(all_pathways_each_pair)) {
  cat("Pathways for interaction", interaction, ":\n")
  print(all_pathways_each_pair[[interaction]])
  cat("\n")
}

# Find unique pathways across all interaction pairs
all_unique_pathways <- unique(unlist(all_pathways_each_pair))

# Print all unique pathways across all interaction pairs
cat("All unique pathways across all interaction pairs:\n")
print(all_unique_pathways)

```

Violin plots
```{r eval=FALSE, include=FALSE}
# Directory to save the results
output_dir <- "~/rocker/RDEB/thesis/violin_plots/"

# Function to plot and save gene expression for a given pathway
plot_and_save <- function(cellchat_object, signaling, tissue, output_dir) {
  plot_obj <- plotGeneExpression(cellchat_object, signaling = signaling)
  plot(plot_obj)
  file_name <- file.path(output_dir, paste0(signaling, "_", tissue, ".png"))
  ggsave(file_name, plot = plot_obj, width = 10, height = 10)
}

# Plot and save gene expression for each common pathway in SCC and skin
for (pathway in all_unique_pathways) {
  plot_and_save(cellchat.t, signaling = pathway, tissue = "tumor", output_dir)
  plot_and_save(cellchat.s, signaling = pathway, tissue = "skin", output_dir)
}
```

Here I create new graphs for abs(Interaction_Difference_abs) top10 interactions.
```{r}
# Filter to get the top 10 interactions by absolute difference
top_10_diff <- merged_df.interactions%>%
  dplyr::ungroup() %>% # Calculate differences without grouping by 'source'
  dplyr::arrange(-abs(Interaction_Difference_abs)) %>%
  dplyr::slice_head(n = 10)

print(top_10_diff)

# Plot top 10 interactions with biggest differences 
difference_plot <- ggplot(
  top_10_diff, 
  aes(
    x = paste(source, target, sep = "-"), 
    y = Interaction_Difference_abs,
    fill = Interaction_Difference_abs > 0)
  ) +
  geom_bar(
    stat = "identity",
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255),
      "FALSE" = rgb(33/255, 102/255, 172/255)
      ),
                    labels = c(
                      "TRUE" = "Tumor",
                      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(
    title = "Top 10 according to Interaction_Difference_abs (Tumor - Skin)", 
    y = "Difference in Interactions", 
    x = "Interaction (Source-target)") +
  theme(axis.text.x = element_text(
    angle = 90, 
    hjust = 1)
    )

print(difference_plot)

ggsave("~/rocker/RDEB/thesis/Top_10_Interaction_Difference_abs.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)

# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, 
           target == current_target)

  # Plot
  p <- ggplot(
    current_data, 
    aes(
      x = ligand, 
      y = receptor, 
      fill = prob.delta)
    ) +
    geom_tile() +
    scale_fill_gradient2(
      low = rgb(33/255, 102/255, 172/255), 
      high = rgb(178/255, 24/255, 43/255), 
      mid = "grey", 
      midpoint = 0,
      limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                 max(current_data$prob.delta, na.rm = TRUE)),
      name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for",
                       current_source,
                       "to",
                       current_target),
         x = "Ligand",
         y = "Receptor"
         ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
      angle = 90, 
      vjust = 0.5
      ),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("~/rocker/RDEB/thesis/interaction_plots_abs/",
                      current_source,
                      "_to_",
                      current_target, 
                      "_abs_",
                      ".pdf"
                      )
  ggsave(
    file_name,
    plot = p,
    width = 10,
    height = 10
    )
  
  # Extracting the relevant interactions from merged_data for the current interaction pair
  filtered_data <- current_data %>%
    distinct(pathway_name)  # Getting unique pathway names involved in these interactions

  # Print the pathways involved in the current interaction pair
  cat("Pathways for interaction", current_source, "to", current_target, ":\n")
  print(filtered_data)

  # Check if these pathways are in the provided pathways lists
  pathways_involved_t <- filtered_data$pathway_name %in% pathways.show.t
  pathways_involved_s <- filtered_data$pathway_name %in% pathways.show.s

  # Print pathways involved both in tumor and skin for the current interaction pair
  cat("Pathways in Tumor:\n")
  print(filtered_data$pathway_name[pathways_involved_t])
  cat("Pathways in Skin:\n")
  print(filtered_data$pathway_name[pathways_involved_s])
}
```
