---
title: "bachelor thesis"
author: "0juliat0"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Libraries
```{r, include=FALSE}
library(monocle3)
library(ggplot2)
library(CellChat)
library(dplyr)
library(tidyr)
library(Matrix)
library(ggplot2)
```

Load data
```{r}
file_path_t <- "/Users/juliat/Desktop/bachelor/Final thesis/ccc_cellchat.t.2.rds"
cellchat.t <- readRDS(file_path_t)
file_path_s <- "/Users/juliat/Desktop/bachelor/Final thesis/ccc_cellchat.s.2.rds"
cellchat.s <- readRDS(file_path_s)
```

Creating dgCMatrix (to avoid turning all data into a matrix, we take ) and merge data 
```{r}
cellchat.t.sC <- subsetCommunication(cellchat.t, thresh=1) # tresh = 1 not to loose data, because it is not possible to compare 13008 and 12058 rows
cellchat.s.sC <- subsetCommunication(cellchat.s, thresh=1)

# merge the two data frames
merged_data <- merge(cellchat.t.sC, cellchat.s.sC, 
                     by = c("source", "target", "ligand", "receptor", "interaction_name",
                            "interaction_name_2", "pathway_name", "annotation", "evidence"),
                     all = TRUE)
```

Replace NA values with 0
```{r}
merged_data <- merged_data %>%
  mutate(
    prob.t = ifelse(is.na(prob.x), 0, prob.x),
    prob.s = ifelse(is.na(prob.y), 0, prob.y),
    pval.t = ifelse(is.na(pval.x), 0, pval.x),
    pval.s = ifelse(is.na(pval.y), 0, pval.y)
  ) %>%
  select(-prob.x, -prob.y, -pval.x, -pval.y) # delete columns 
```

Calculating the difference between probabilities tumor - skin
```{r}
merged_data <- merged_data %>%
  mutate(prob.delta = prob.t - prob.s) %>%
  arrange(desc(abs(prob.delta)))
```

Count interactions for skin and tumor (actual numeric difference )
```{r}
cellchat.t.sC_counts <- cellchat.t.sC %>% group_by(source, target) %>% summarise(count_t = n())
cellchat.s.sC_counts <- cellchat.s.sC %>% group_by(source, target) %>% summarise(count_s = n())
# Merge counts
merged_df.interactions <- merge(cellchat.t.sC_counts, cellchat.s.sC_counts, by=c("source", "target"), all = TRUE)
# Replacing NA with 0 in case some interactions are not present in one of the data frames
merged_df.interactions[is.na(merged_df.interactions)] <- 0
# Calculate the difference in counts:
merged_df.interactions$interaction_difference <- merged_df.interactions$count_t - merged_df.interactions$count_s
```

Heatmap based on counts difference
```{r}
heatmap_plot <- ggplot(merged_df.interactions, aes(x=target, y=source, fill=interaction_difference)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Difference in Number of Interactions (Tumor - Skin)", x = "target (receptor)", y = "source (ligand)", fill = "Interactions Difference") +
  scale_fill_gradient2(name = "Interactions Diff", high = rgb(178/255, 24/255, 43/255), low = rgb(33/255, 102/255, 172/255)) # Diverging color scale
print(heatmap_plot)
```

Bubble Plot
```{r}
netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 8, thresh = 0.05) # Fibroblast_32 to Fibroblast_32
netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 28, thresh = 0.05) # Fibroblast_32 to T naive CD4
netVisual_bubble(cellchat.t, sources.use = 3, targets.use = 28, thresh = 0.05) # Endothel_vasc._31 to T naive CD4 
netVisual_bubble(cellchat.t, sources.use = 9, targets.use = 9, thresh = 0.05) # Fibroblast_36 to Fibroblast_36
netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 2, thresh = 0.05) # Fibroblast_32 to Endothel
netVisual_bubble(cellchat.t, sources.use = 3, targets.use = 8, thresh = 0.05) # Endothel_vasc._31 to Fibroblast_32 
netVisual_bubble(cellchat.t, sources.use = 5, targets.use = 28, thresh = 0.05) # Fibroblast_18 to T naive CD4  5 28
netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 13, thresh = 0.05) # Fibroblast_32 to Monocyte 
netVisual_bubble(cellchat.t, sources.use = 9, targets.use = 28, thresh = 0.05) # Fibroblast_36 to T naive CD4
netVisual_bubble(cellchat.t, sources.use = 6, targets.use = 2, thresh = 0.05) # Fibroblast_20 to Endothel
```

CELLCHAT tumor pathways
```{r}
# cellchat.t@netP$pathways
# cellchat.t@meta$DE = factor(cellchat.t@meta$DE, levels = c ("B cell", "Endothel", "Endothel_vasc._31", "Erythrocyte", "Fibroblast_18", "Fibroblast_20", "Fibroblast_25", "Fibroblast_32", "Fibroblast_36", "T gd", "Keratinocyte_26", "Mast_27", "Monocyte", "Monocyte_23", "Neurons_47", "Neutrophil", "NK", "NK T like", "Plasma_15", "Plasma_35", "Plasma_37", "Platelet", "Schwann_43", "T cytotox. CD8", "T fh", "T mem 1", "T mem 2", "T naive CD4", "T reg CD4", "T undef."))
# x <- "COLLAGEN"
# for(x in cellchat.t@netP$pathways){
#   p <- plotGeneExpression(cellchat.t, signaling = x)
#   ggsave(paste0(x, ".pdf"),plot=p, width=10, height=10)
#   }
```

CELLCHAT skin pathways
```{r}
# cellchat.s@netP$pathways
# cellchat.s@meta$DE = factor(cellchat.s@meta$DE, levels = c ("B cell", "Endothel", "Endothel_vasc._31", "Erythrocyte", "Fibroblast_18", "Fibroblast_20", "Fibroblast_25", "Fibroblast_32", "Fibroblast_36", "T gd", "Keratinocyte_26", "Mast_27", "Monocyte", "Monocyte_23", "Neurons_47", "Neutrophil", "NK", "NK T like", "Plasma_15", "Plasma_35", "Plasma_37", "Platelet", "Schwann_43", "T cytotox. CD8", "T fh", "T mem 1", "T mem 2", "T naive CD4", "T reg CD4", "T undef."))
# y <- "COLLAGEN"
# for(y in cellchat.s@netP$pathways){
#   p <- plotGeneExpression(cellchat.s, signaling = y)
#   ggsave(paste0(y, ".pdf"),plot=p, width=10, height=10)
# }
```
Number of interactions per source cell (ligand) in tumor and skin
```{r}
df.t <- cellchat.t %>% subsetCommunication(thresh = NA) %>% as_tibble() # creating a tibble
df.s <- cellchat.s %>% subsetCommunication(thresh = NA) %>% as_tibble()

df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>%
  dplyr::count(source) %>%
  dplyr::rename(Number_of_Interactions = n)

df.t$Type <- "Tumor"

df.s <- cellchat.s %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)

df.s$Number_of_Interactions <- -df.s$Number_of_Interactions # making counts negative
df.s$Type <- "Skin"

combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin
combined_plot <- ggplot(combined_data, aes(x = source, y = Number_of_Interactions, fill = Type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Number of interactions per source cell", y = "Number of Interactions", x = "Source Cell") +
  scale_fill_manual(values = c("Skin" = rgb(33/255, 102/255, 172/255), "Tumor" = rgb(178/255, 24/255, 43/255))) +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1) # hjust = 1 to "press" names to plot
  )
print(combined_plot)

```
There are no T naive CD4 in skin so lets exclude them from analysis 
```{r}

# Define the identifier for naive CD4 T cells
naive_CD4_identifier <- "T naive CD4"

# Filter out interactions involving naive CD4 T cells from the tumor dataset
cellchat.t.sC <- subsetCommunication(cellchat.t, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | target == naive_CD4_identifier))

# Filter out interactions involving naive CD4 T cells from the skin dataset
cellchat.s.sC <- subsetCommunication(cellchat.s, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | target == naive_CD4_identifier))

# merge the two data frames
merged_data <- merge(cellchat.t.sC, cellchat.s.sC, 
                     by = c("source", "target", "ligand", "receptor", "interaction_name",
                            "interaction_name_2", "pathway_name", "annotation", "evidence"),
                     all = TRUE)

```
Repeat all steps bellow but for data without T naive CD4
Replace NA values with 0
```{r}
merged_data <- merged_data %>%
  mutate(
    prob.t = ifelse(is.na(prob.x), 0, prob.x),
    prob.s = ifelse(is.na(prob.y), 0, prob.y),
    pval.t = ifelse(is.na(pval.x), 0, pval.x),
    pval.s = ifelse(is.na(pval.y), 0, pval.y)
  ) %>%
  select(-prob.x, -prob.y, -pval.x, -pval.y)
```

Calculating the difference between probabilities tumor - skin
```{r}
merged_data <- merged_data %>%
  mutate(prob.delta = prob.t - prob.s) %>%
  arrange(desc(abs(prob.delta)))
```

Count interactions for skin and tumor (actual numeric difference )
```{r}
cellchat.t.sC_counts <- cellchat.t.sC %>% group_by(source, target) %>% summarise(count_t = n())
cellchat.s.sC_counts <- cellchat.s.sC %>% group_by(source, target) %>% summarise(count_s = n())
# Merge counts
merged_df.interactions <- merge(cellchat.t.sC_counts, cellchat.s.sC_counts, by=c("source", "target"), all = TRUE)
# Replacing NA with 0 in case some interactions are not present in one of the data frames
merged_df.interactions[is.na(merged_df.interactions)] <- 0
# Calculate the difference in counts:
merged_df.interactions$interaction_difference <- merged_df.interactions$count_t - merged_df.interactions$count_s
```

Heatmap based on counts difference
```{r}
heatmap_plot <- ggplot(merged_df.interactions, aes(x=target, y=source, fill=interaction_difference)) + 
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Difference in Number of Interactions (Tumor - Skin)", x = "target (receptor)", y = "source (ligand)", fill = "Interactions Difference") +
  scale_fill_gradient2(name = "Interactions Diff", high = rgb(178/255, 24/255, 43/255), low = rgb(33/255, 102/255, 172/255)) # Diverging color scale
print(heatmap_plot)
```

Number of interactions per source cell (ligand) in tumor and skin
```{r}
# Count interactions per source for tumor data without T naive CD4
df.t <- merged_data %>%
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)
df.t$Type <- "Tumor"

# Count interactions per source for skin data without T naive CD4
df.s <- merged_data %>%
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)
df.s$Number_of_Interactions <- -df.s$Number_of_Interactions # making counts negative
df.s$Type <- "Skin"

# Combine the two datasets
combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin without T naive CD4
combined_plot <- ggplot(combined_data, aes(x = source, y = Number_of_Interactions, fill = Type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Number of interactions per source cell without T naive CD4", y = "Number of Interactions", x = "Source Cell") +
  scale_fill_manual(values = c("Skin" = rgb(33/255, 102/255, 172/255), "Tumor" = rgb(178/255, 24/255, 43/255))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(combined_plot)


```

Number of interactions per source cell (ligand) in tumor - skin
```{r}
# Compute interaction counts for each dataset, excluding T naive CD4
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>%
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% # no need to exclude, since in skin there is nothing 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'source' (28 levels - celltypes) column
combined_data.diff <- full_join(df.t, df.s, by = "source")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[is.na(combined_data.diff$Tumor_Interactions)] <- 0
combined_data.diff$Skin_Interactions[is.na(combined_data.diff$Skin_Interactions)] <- 0

# Compute the difference for each source: skin - tumor
combined_data.diff$Interaction_Difference <- combined_data.diff$Tumor_Interactions - combined_data.diff$Skin_Interactions

# Display the table (screenshot saved)
print(combined_data.diff, n = 27)# Display the table (screenshot saved)

# Plot tumor - skin with colors for positive and negative differences
difference_plot <- ggplot(combined_data.diff, aes(x = source, y = Interaction_Difference, 
                                                   fill = Interaction_Difference > 0)) + # Moved inside aes()
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("TRUE" = rgb(178/255, 24/255, 43/255), "FALSE" = rgb(33/255, 102/255, 172/255)),
                    labels = c("TRUE" = "Tumor", "FALSE" = "Skin")) +
  theme_minimal() +
  labs(title = "Difference in interactions per source cell (Tumor - Skin)", 
       y = "Difference in Interactions", 
       x = "Source Cell") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

difference_plot

```


Scatter plot source
```{r}
scatter_plot <- ggplot(combined_data.diff, aes(x = Tumor_Interactions, y = Skin_Interactions, label = source)) +
  geom_point(aes(color = source), size = 3) +
  geom_text(aes(label = source), size = 3) +
  geom_smooth(method = "lm",se = TRUE) +
  labs(title = "Number of Interactions per source Cell", x = "Tumor Interactions", y = "Skin Interactions") +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)
```

Number of interactions per target cell (receptor) in tumor and skin
```{r}
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>%
  dplyr::count(target) %>%
  dplyr::rename(Number_of_Interactions = n)

df.t$Type <- "Tumor"

df.s <- cellchat.s %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Number_of_Interactions = n)

df.s$Number_of_Interactions <- -df.s$Number_of_Interactions # making counts negative
df.s$Type <- "Skin"

combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin
combined_plot <- ggplot(combined_data, aes(x = target, y = Number_of_Interactions, fill = Type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Number of interactions per target cell", y = "Number of Interactions", x = "Target Cell") +
  scale_fill_manual(values = c("Skin" = rgb(33/255, 102/255, 172/255), "Tumor" = rgb(178/255, 24/255, 43/255))) +
    theme_minimal()  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1) # hjust = 1 to "press" names to plot
  )
print(combined_plot)

```

Number of interactions per target cell (receptor) in tumor - skin
```{r}
# Compute interaction counts for each dataset
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'target' (28 levels - celltypes) column
combined_data.diff <- full_join(df.t, df.s, by = "target")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[is.na(combined_data.diff$Tumor_Interactions)] <- 0
combined_data.diff$Skin_Interactions[is.na(combined_data.diff$Skin_Interactions)] <- 0

# Compute the difference for each target: tumor - skin
combined_data.diff$Interaction_Difference <- combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions
print (combined_data.diff, n = 27) # Display the table (screenshot saved)
# Plot tumor - skin with colors for positive and negative differences
difference_plot <- ggplot(combined_data.diff, aes(x = target, y = Interaction_Difference, fill = Interaction_Difference > 0)) + # Moved inside aes()
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("TRUE" = rgb(178/255, 24/255, 43/255), "FALSE" = rgb(33/255, 102/255, 172/255)),
                    labels = c("TRUE" = "Tumor", "FALSE" = "Skin")) +
  theme_minimal() +
  labs(title = "Difference in interactions per target cell (Tumor - Skin)", 
       y = "Difference in Interactions", 
       x = "Target Cell") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

difference_plot
```

Scatter plot target
```{r}
scatter_plot <- ggplot(combined_data.diff, aes(x = Tumor_Interactions, y = Skin_Interactions, label = target)) + 
  geom_point(aes(color = target), size = 3) +
  geom_text(aes(label = target), size = 3) +
  geom_smooth(method = "lm",se = TRUE) +
  labs(title = "Number of Interactions per target Cell", x = "Tumor Interactions", y = "Skin Interactions") +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)
```

Plotting top 10 interactions with biggest differences in tumor - skin
```{r}
# Combining data tumor and skin 
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>%
  dplyr::summarise(Tumor_Interactions = n())

df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>% 
  dplyr::summarise(Skin_Interactions = n())

# Merge datasets by the 'source' and 'target'
combined_data.diff <- full_join(df.t, df.s, by = c("source", "target"))

# Replace NAs with 0
combined_data.diff$Tumor_Interactions[is.na(combined_data.diff$Tumor_Interactions)] <- 0
combined_data.diff$Skin_Interactions[is.na(combined_data.diff$Skin_Interactions)] <- 0

# Compute the difference for each source-target pair: skin - tumor
combined_data.diff$Interaction_Difference <- combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions

# Filter to get the top 10 interactions by absolute difference
top_10_diff <- combined_data.diff %>%
  dplyr::ungroup() %>% # Calculate differences without grouping by 'source'
  dplyr::arrange(-abs(Interaction_Difference)) %>%
  dplyr::slice_head(n = 10)

print(top_10_diff)

# paired t-test
result <- t.test(top_10_diff$Tumor_Interactions, top_10_diff$Skin_Interactions, paired=TRUE)
print(result)
###p-value = 0.004082 => statistically significant difference between "Tumor_Interactions" and "Skin_Interactions"

# Plot top 10 interactions with biggest differences 
difference_plot <- ggplot(top_10_diff, aes(x = paste(source, target, sep = "-"), y = Interaction_Difference,fill = Interaction_Difference > 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("TRUE" = rgb(178/255, 24/255, 43/255), "FALSE" = rgb(33/255, 102/255, 172/255)),
                    labels = c("TRUE" = "Tumor", "FALSE" = "Skin")) +
  theme_minimal() +
  labs(title = "Top 10 Differences in Interactions (Tumor - Skin)", y = "Difference in Interactions", x = "Interaction (Source-target)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(difference_plot)


```

Function to process interactions between sorce and target celltypes 
```{r}
plot_interaction <- function(df, source_value, target_value) {
  df_filtered <- df %>%
    filter(source == source_value & target == target_value) %>%
    slice_max(order_by = abs(prob.delta), n = 100) %>%
    select(-c(pval.t, pval.s))

  # Create a label from the source and target values
  interaction_label <- paste(source_value, target_value, sep = " to ")

  # Pivot and plot using the new label
  df_filtered %>%
    pivot_longer(cols = c("prob.s", "prob.t"), names_to = "variable", values_to = "value") %>%
    ggplot(aes(x = interaction_name_2, y = value, fill = variable)) +
    geom_col(position = "dodge") +
    scale_fill_manual(values = c("prob.s" = rgb(33/255, 102/255, 172/255), "prob.t" = rgb(178/255, 24/255, 43/255))) +
    labs(x = interaction_label, y = "Probability", fill = "Probability") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
}
```

Plotting top 10 interactions based on abs(interactions difference)
```{r}
plot_interaction(merged_data, "Fibroblast_32", "Fibroblast_32") +
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_36", "Fibroblast_36")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_32", "Endothel")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Endothel_vasc._31", "Fibroblast_32")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_32", "Monocyte")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_20", "Endothel")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_32", "Fibroblast_20")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_18", "Endothel")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Fibroblast_18", "Fibroblast_36")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot_interaction(merged_data, "Endothel", "Endothel")+
    theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Sorting interactions in tumor based on pval and prob
```{r}
# Identifying the top interactions based on the p-value (interactions with pval == 0)
# Group by interaction_name, then filter where pval in the group are 0
interactions_pval_zero.t <- cellchat.t.sC %>%
  group_by(interaction_name) %>%
  filter(pval == 0)

  
# Calculate the number of interactions to select for the top 1%
top_1_percent_count <- ceiling(nrow(interactions_pval_zero.t) * 0.01) # Use ceiling() to round up

# Create a new dataframe sorted by prob and select top 1%
interactions_sorted_prob.t <- interactions_pval_zero.t %>%
  ungroup() %>% # Ungroup to ensure slice_max works as expected
  slice_max(order_by = prob, n = top_1_percent_count)
  
# all interactions summarised, so each interaction = row
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
ggplot(interactions_sorted_prob.t, aes(x = source, y = target, label = interaction_name)) +
  geom_tile(aes(fill = interaction_name), colour = "white") +
  labs(title = "Top 1% Interactions in Tumor", x = "Source", y = "Target", fill = "Interactions") +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )
  
# Retrieve the full data for these top interactions
details_top_interactions_t <- cellchat.t.sC %>%
  filter(interaction_name %in% interactions_sorted_prob.t$interaction_name)
#  interactions are NOT summarised, so each interaction = multiple rows
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
ggplot(details_top_interactions_t, aes(x = source, y = target, label = interaction_name)) +
  geom_tile(aes(fill = interaction_name), colour = "white") +
  labs(title = "Top 1% Interactions in Tumor", x = "Source", y = "Target", fill = "Interactions") +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )
```

Sorting interactions in skin based on pval and prob
```{r}
# Identifying the top interactions based on the p-value (interactions with pval == 0)
# Group by interaction_name, then filter where pval in the group are 0
interactions_pval_zero.s <- cellchat.s.sC %>%
  group_by(interaction_name) %>%
  filter(pval == 0)
  
# Calculate the number of interactions to select for the top 1%
top_1_percent_count <- ceiling(nrow(interactions_pval_zero.s) * 0.01) # Use ceiling to round up
# Create a new dataframe sorted by prob and select top 1%
interactions_sorted_prob.s <- interactions_pval_zero.s %>%
  ungroup() %>% # Ungroup to ensure slice_max works as expected
  slice_max(order_by = prob, n = top_1_percent_count)
  
# all interactions summarised, so each interaction = row
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
ggplot(interactions_sorted_prob.s, aes(x = source, y = target, label = interaction_name)) +
  geom_tile(aes(fill = interaction_name), colour = "white") +
  labs(title = "Top 1% Interactions in Skin", x = "Source", y = "Target", fill = "Interactions") +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )
  
# Retrieve the full data for these top interactions
details_top_interactions_s <- cellchat.s.sC %>%
  filter(interaction_name %in% interactions_sorted_prob.s$interaction_name)
#  interactions are NOT summarised, so each interaction = multiple rows
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
ggplot(details_top_interactions_s, aes(x = source, y = target, label = interaction_name)) +
  geom_tile(aes(fill = interaction_name), colour = "white") +
  labs(title = "Top 1% Interactions in Skin", x = "Source", y = "Target", fill = "Interactions") +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )


```
Top 100 Interactions with Biggest Difference in Probability Tumor
```{r}
# Filtering by pval == 0 and abs(prob.delta)
merged_common_data <- merged_data %>%
  filter(pval.t == 0 & pval.s == 0) %>%
  mutate(
    more_common_in = ifelse(prob.delta > 0, "Tumor", "Skin")
  )%>%
  slice_max(order_by = abs(prob.delta), n = 100)

# Plotting
ggplot(merged_common_data, aes(x = source, y = target)) +
  geom_tile(aes(fill = more_common_in), colour = "white") +
  geom_text(aes(label = interaction_name), size = 3, check_overlap = TRUE, vjust = "top") + # Add interaction names
  scale_fill_manual(values = c("Tumor" = rgb(178/255, 24/255, 43/255), "Skin" = rgb(33/255, 102/255, 172/255))) + # Define custom colors for tumor and skin
  labs(title = "Top 100 Interactions with Biggest Difference in Probability (Tumor vs. Skin)", 
       x = "Source", y = "Target", fill = "More Common In") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```

Interaction based on cell types
```{r}
# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, target == current_target)

  # Plot
  p <- ggplot(current_data, aes(x = ligand, y = receptor, fill = prob.delta)) +
    geom_tile() +
    scale_fill_gradient2(low = rgb(33/255, 102/255, 172/255), high = rgb(178/255, 24/255, 43/255), mid = "grey", midpoint = 0,
                         limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                                    max(current_data$prob.delta, na.rm = TRUE)),
                         name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for", current_source, "to", current_target),
         x = "Ligand", y = "Receptor") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("/Users/juliat/Desktop/bachelor/", current_source, "_to_", current_target, ".pdf")
  ggsave(file_name, plot = p, width = 10, height = 10)
}
```

Violing plots
```{r}
# COLLAGEN.t <- plotGeneExpression(cellchat.t, signaling = c("COLLAGEN"))
# plot(COLLAGEN.t)
# COLLAGEN.s <- plotGeneExpression(cellchat.s, signaling = c("COLLAGEN"))
# plot(COLLAGEN.s)
# LAMININ.t <- plotGeneExpression(cellchat.t, signaling = c("LAMININ"))
# plot(LAMININ.t)
# LAMININ.s <- plotGeneExpression(cellchat.s, signaling = c("LAMININ"))
# plot(LAMININ.s)
# THBS.t <- plotGeneExpression(cellchat.t, signaling = c("THBS"))
# plot(THBS.t)
# THBS.s <- plotGeneExpression(cellchat.s, signaling = c("THBS"))
# plot(THBS.s)
# TENASCIN.t <- plotGeneExpression(cellchat.t, signaling = c("TENASCIN"))
# plot(TENASCIN.t)
# TENASCIN.s <- plotGeneExpression(cellchat.s, signaling = c("TENASCIN"))
# plot(TENASCIN.s)
# SEMA3.t <- plotGeneExpression(cellchat.t, signaling = c("SEMA3"))
# plot(SEMA3.t)
# SEMA3.s <- plotGeneExpression(cellchat.s, signaling = c("SEMA3"))
# plot(SEMA3.s)
# SELE.t <- plotGeneExpression(cellchat.t, signaling = c("SELE"))
# plot(SELE.t)
# SELE.s <- plotGeneExpression(cellchat.s, signaling = c("SELE"))
# plot(SELE.s)
# NEGR.t <- plotGeneExpression(cellchat.t, signaling = c("NEGR"))
# plot(NEGR.t)
# NEGR.s <- plotGeneExpression(cellchat.s, signaling = c("NEGR"))
# plot(NEGR.s)
# MHCII.t <- plotGeneExpression(cellchat.t, signaling = c("MHC-II"))
# plot(MHCII.t)
# MHCII.s <- plotGeneExpression(cellchat.s, signaling = c("MHC-II"))
# plot(MHCII.s)
# MHCI.t <- plotGeneExpression(cellchat.t, signaling = c("MHC-I"))
# plot(MHCI.t)
# MHCI.s <- plotGeneExpression(cellchat.s, signaling = c("MHC-I"))
# plot(MHCI.s)
# VISFATIN.t <- plotGeneExpression(cellchat.t, signaling = c("VISFATIN"))
# plot(VISFATIN.t)
# VISFATIN.s <- plotGeneExpression(cellchat.s, signaling = c("VISFATIN"))
# plot(VISFATIN.s)
```
  