---
title: "bachelor thesis"
author: "0juliat0"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r, include=FALSE}
library(plotly)
library(ggplot2)
library(CellChat)
library(dplyr)
library(tidyr)
library(Matrix)
library(ggplot2)
library(pheatmap)
library(Seurat)
```

Load data
```{r}
file_path_t <- "~/rocker/RDEB/ccc_cellchat.t.2.rds"
cellchat.t <- readRDS(file_path_t)
file_path_s <- "~/rocker/RDEB/ccc_cellchat.s.2.rds"
cellchat.s <- readRDS(file_path_s)
```

Creating dgCMatrix (to avoid turning all data into a matrix, we take ) and merge data
```{r}
# tresh = 1 not to loose data, because it is not possible to compare 13008 and 12058 rows
cellchat.t.sC <- subsetCommunication(cellchat.t, thresh=1) 
cellchat.s.sC <- subsetCommunication(cellchat.s, thresh=1)

# Merge the two data frames
merged_data <- merge(cellchat.t.sC,
                     cellchat.s.sC, 
                     by = c("source", "target", "ligand", "receptor", 
                            "interaction_name","interaction_name_2", 
                            "pathway_name", "annotation", "evidence"),
                     all = TRUE)
```

Replace NA values with 0
```{r}
merged_data <- merged_data %>%
  mutate(
    prob.t = ifelse(is.na(prob.x), 0, prob.x),
    prob.s = ifelse(is.na(prob.y), 0, prob.y),
    pval.t = ifelse(is.na(pval.x), 0, pval.x),
    pval.s = ifelse(is.na(pval.y), 0, pval.y)
  ) %>%
  select(-prob.x, -prob.y, -pval.x, -pval.y) # delete columns 
```

Calculating the difference between probabilities tumor - skin
```{r}
merged_data <- merged_data %>%
  mutate(prob.delta = prob.t - prob.s) %>%
  arrange(desc(abs(prob.delta)))
```

Count interactions for skin and tumor (actual numeric difference )
```{r}
cellchat.t.sC_counts <-
  cellchat.t.sC %>%
  group_by(source, target) %>%
  summarise(count_t = n())

cellchat.s.sC_counts <- 
  cellchat.s.sC %>% 
  summarise(.by = c(source, target), count_s = n())

# Merge counts
merged_df.interactions <- merge(
  cellchat.t.sC_counts,
  cellchat.s.sC_counts,
  .by = c("source", "target"),
  all = TRUE
)
# Replacing NA with 0 in case some interactions are not present in one of the data frames
merged_df.interactions[is.na(merged_df.interactions)] <- 0
# Calculate the difference in counts:
merged_df.interactions$interaction_difference <-
  merged_df.interactions$count_t - merged_df.interactions$count_s
```

Heatmap based on counts difference
```{r}
heatmap_plot <-
  ggplot(merged_df.interactions,
         aes(x=target, y=source, fill=interaction_difference)) + 
  geom_tile(color = "white", linewidth = 0.1) +
  scale_fill_gradient2(
    name = "Interactions Difference", 
    high = rgb(178/255, 24/255, 43/255), # Diverging color scale
    low = rgb(33/255, 102/255, 172/255)
  ) +
  labs(
    title = "Difference in Number of Interactions (Tumor vs Skin)",
    x = "Target cells (receptor)",
    y = "Source cells (ligand)",
    fill = "Interactions Difference"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
   
print(heatmap_plot)
ggsave("~/rocker/RDEB/thesis/heatmap_based_on_counts_difference.png", 
       plot = heatmap_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)

# Convert to plotly for interactivity
interactive_heatmap <- ggplotly(heatmap_plot)

# Print the interactive heatmap
interactive_heatmap
```
```{r}
# Convert factors to characters to avoid conversion issues
merged_df.interactions$source <- as.character(merged_df.interactions$source)
merged_df.interactions$target <- as.character(merged_df.interactions$target)

# Create a new matrix with all zero values
interaction_matrix <- matrix(0, 
                             nrow = length(unique(merged_df.interactions$source)), 
                             ncol = length(unique(merged_df.interactions$target)),
                             dimnames = list(unique(merged_df.interactions$source),
                             unique(merged_df.interactions$target)))

# Fill the matrix with interaction_difference values
for(i in seq_len(nrow(merged_df.interactions))) {
  interaction_matrix[merged_df.interactions$source[i],
                     merged_df.interactions$target[i]] <- 
    merged_df.interactions$interaction_difference[i]
}

# Pheatmap
pheatmap_based_on_counts_difference <- pheatmap(
  interaction_matrix,
         color = colorRampPalette(
           c(rgb(33/255, 102/255, 172/255),
             "white",
             rgb(178/255, 24/255, 43/255)))(100),
  border_color = NA,
  treeheight_row = 50,
  treeheight_col = 50,
  legend = TRUE,
  main = "Difference in Number of Interactions (Tumor vs Skin)"
         )

print (pheatmap_based_on_counts_difference)

ggsave("~/rocker/RDEB/thesis/pheatmap_based_on_counts_difference.png", 
       plot = pheatmap_based_on_counts_difference, 
       width = 10, 
       height = 8, 
       dpi = 300)
                         
```
Bubble Plot
```{r}
# netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 8, thresh = 0.05) # Fibroblast_32 to Fibroblast_32
# netVisual_bubble(cellchat.t, sources.use = 9, targets.use = 9, thresh = 0.05) # Fibroblast_36 to Fibroblast_36
# netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 2, thresh = 0.05) # Fibroblast_32 to Endothel
# netVisual_bubble(cellchat.t, sources.use = 3, targets.use = 8, thresh = 0.05) # Endothel_vasc._31 to Fibroblast_32 
# netVisual_bubble(cellchat.t, sources.use = 8, targets.use = 13, thresh = 0.05) # Fibroblast_32 to Monocyte
# netVisual_bubble(cellchat.t, sources.use = 6, targets.use = 2, thresh = 0.05) # Fibroblast_20 to Endothel
```

Number of interactions per source cell (ligand) in tumor and skin
```{r}
# Creating a tibble
df.t <- cellchat.t %>% 
  subsetCommunication(thresh = NA) %>% 
  as_tibble() 
df.s <- cellchat.s %>%
  subsetCommunication(thresh = NA) %>%
  as_tibble()

df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>%
  dplyr::count(source) %>%
  dplyr::rename(Number_of_Interactions = n)

df.t$Type <- "Tumor"

df.s <- cellchat.s %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)

# Making counts negative
df.s$Number_of_Interactions <- -df.s$Number_of_Interactions 
df.s$Type <- "Skin"

combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin
combined_plot <- ggplot(
  combined_data, 
  aes(
    x = source, 
    y = Number_of_Interactions, 
    fill = Type)
  ) +
  geom_bar(
    stat = "identity", 
    position = "stack") +
  labs(
    title = "Number of interactions per source cell", 
    y = "Number of Interactions", 
    x = "Source Cell") +
  scale_fill_manual(
    values = c(
      "Skin" = rgb(33/255, 102/255, 172/255), 
      "Tumor" = rgb(178/255, 24/255, 43/255))
    ) +
    theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1) 
    # hjust = 1 to "press" names to plot
  )
print(combined_plot)

ggsave("~/rocker/RDEB/thesis/count_source_tumor_vs_skin.png", 
       plot = combined_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)

```
There are no T naive CD4 in skin so lets exclude them from analysis 
```{r}
# Define the identifier for naive CD4 T cells
naive_CD4_identifier <- "T naive CD4"

# Filter out interactions involving naive CD4 T cells from the tumor dataset
cellchat.t.sC <- subsetCommunication(cellchat.t, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | 
                            target == naive_CD4_identifier))

# Filter out interactions involving naive CD4 T cells from the skin dataset
cellchat.s.sC <- subsetCommunication(cellchat.s, thresh=1) %>%
                 filter(!(source == naive_CD4_identifier | 
                            target == naive_CD4_identifier))

# merge the two data frames
merged_data <- merge(cellchat.t.sC, cellchat.s.sC, 
                     by = c("source", "target", "ligand", "receptor", 
                            "interaction_name","interaction_name_2", 
                            "pathway_name", "annotation", "evidence"),
                     all = TRUE)

```

Repeat all steps bellow but for data without T naive CD4
Replace NA values with 0
```{r}
merged_data <- merged_data %>%
  mutate(
    prob.t = ifelse(is.na(prob.x), 0, prob.x),
    prob.s = ifelse(is.na(prob.y), 0, prob.y),
    pval.t = ifelse(is.na(pval.x), 0, pval.x),
    pval.s = ifelse(is.na(pval.y), 0, pval.y)
  ) %>%
  select(-prob.x, -prob.y, -pval.x, -pval.y)
```

Calculating the difference between probabilities tumor - skin
```{r}
merged_data <- merged_data %>%
  mutate(prob.delta = prob.t - prob.s) %>%
  arrange(desc(abs(prob.delta)))
```

Count interactions for skin and tumor (actual numeric difference )
```{r}
# Create the summary data frame with the desired columns
merged_df.interactions <- merged_data %>%
  group_by(source, target) %>%
  summarise(
    Tumor_Interactions = sum(prob.t > 0),
    Skin_Interactions = sum(prob.s > 0),
    Interaction_Difference = Tumor_Interactions - Skin_Interactions,
    Interaction_Difference_abs = sum(xor(prob.t == 0, prob.s == 0)),
    .groups = 'drop'
  )

# Print the interaction summary data frame
print(merged_df.interactions)
```

Heatmap based on counts difference
```{r}
heatmap_plot <- ggplot(
  merged_df.interactions,
  aes(
    x=target,
    y=source,
    fill=Interaction_Difference)
  ) +
  geom_tile(color = "white", size = 0.1) +
  scale_fill_gradient2(name = "Interactions Difference",
                       high = rgb(178/255, 24/255, 43/255),
                       low = rgb(33/255, 102/255, 172/255)
                       ) + # Diverging color scale
  labs(title = "Difference in Number of Interactions (Tumor vs Skin)",
       x = "Target cells (receptor)",
       y = "Source cells (ligand)",
       fill = "Interactions Difference") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
)

print(heatmap_plot)

ggsave("~/rocker/RDEB/thesis/heatmap_based_on_counts_difference_without_naive_CD4.png",
       plot = heatmap_plot,
       width = 10,
       height = 8,
       dpi = 300)

# Convert to plotly for interactivity
interactive_heatmap <- ggplotly(heatmap_plot)

# Print the interactive heatmap
interactive_heatmap

# Create a point plot with color and size adjustments
point_plot <- ggplot(merged_df.interactions, aes(x = target, y = source)) +
  geom_point(aes(size = Interaction_Difference_abs, color = Interaction_Difference), shape = 20) +
  scale_size_continuous(name = "Absolute Interaction Difference", range = c(1, 11)) + 
  scale_color_gradient2(name = "Interactions Difference",
                        high = rgb(178/255, 24/255, 43/255), 
                        mid = "white",
                        low = rgb(33/255, 102/255, 172/255),
                        midpoint = 0) + # Diverging color scale
  labs(title = "Difference in Number of Interactions (Tumor vs Skin)",
       x = "Target cells (receptor)",
       y = "Source cells (ligand)",
       color = "Interactions Difference",
       size = "Abs Interaction Diff") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right",
    legend.direction = "vertical",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

# Print the point plot
print(point_plot)

# Save the plot
ggsave("~/rocker/RDEB/thesis/point_plot_based_on_interactions_difference.png",
       plot = point_plot,
       width = 10,
       height = 8,
       dpi = 300)
```

```{r}
# Convert factors to characters to avoid conversion issues
merged_df.interactions$source <- as.character(merged_df.interactions$source)
merged_df.interactions$target <- as.character(merged_df.interactions$target)

# Create a new matrix with all zero values
interaction_matrix <- matrix(0, 
                             nrow = length(
                               unique(merged_df.interactions$source)
                               ), 
                             ncol = length(
                               unique(merged_df.interactions$target)
                               ),
                             dimnames = list(
                               unique(merged_df.interactions$source),
                               unique(merged_df.interactions$target)
                               )
                             )

# Fill the matrix with interaction_difference values
for(i in seq_len(nrow(merged_df.interactions))) {
  interaction_matrix [merged_df.interactions$source[i],
                      merged_df.interactions$target[i]] <- 
    merged_df.interactions$Interaction_Difference[i]
}

# Pheatmap
pheatmap_based_on_counts_difference <- pheatmap(
  interaction_matrix,
  color = colorRampPalette(c(rgb(33/255, 102/255, 172/255), 
                             "white", 
                             rgb(178/255, 24/255, 43/255)))(100),
         border_color = NA,
         treeheight_row = 50,
         treeheight_col = 50,
         legend = TRUE,
          main = "Difference in Number of Interactions (Tumor vs Skin)"
         )

print (pheatmap_based_on_counts_difference)

ggsave("~/rocker/RDEB/thesis/pheatmap_based_on_counts_difference_without_CD4.png", 
       plot = pheatmap_based_on_counts_difference, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Number of interactions per source cell (ligand) in tumor and skin
```{r}
# Count interactions per source for tumor data without T naive CD4
df.t <- merged_data %>%
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)
df.t$Type <- "Tumor"

# Count interactions per source for skin data without T naive CD4
df.s <- merged_data %>%
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Number_of_Interactions = n)
# Making counts negative
df.s$Number_of_Interactions <- -df.s$Number_of_Interactions 
df.s$Type <- "Skin"

# Combine the two datasets
combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin without T naive CD4
combined_plot <- ggplot(
  combined_data, 
  aes(x = source, 
      y = Number_of_Interactions, 
      fill = Type)) +
  geom_bar(
    stat = "identity", 
    position = "stack") +
  labs(title = "Number of interactions per source cell without T naive CD4", 
       y = "Number of Interactions", 
       x = "Source Cell") +
  scale_fill_manual(values = c("Skin" = rgb(33/255, 102/255, 172/255), 
                               "Tumor" = rgb(178/255, 24/255, 43/255))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(combined_plot)

ggsave("~/rocker/RDEB/thesis/count_source_tumor_vs_skin_without_naive_CD4.png", 
       plot = combined_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Number of interactions per source cell (ligand) in tumor - skin
```{r}
# Compute interaction counts for each dataset, excluding T naive CD4
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>%
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% # no need to exclude, since in skin there is nothing 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(source) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'source' (27 levels - celltypes) column
combined_data.diff <- full_join(df.t, df.s, by = "source")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each source: skin - tumor
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions - combined_data.diff$Skin_Interactions

# Display the table (screenshot saved)
print(combined_data.diff, n = 27)# Display the table (screenshot saved)

# Plot tumor - skin with colors for positive and negative differences
difference_plot <- ggplot(
  combined_data.diff, 
  aes(
    x = source, 
    y = Interaction_Difference, 
    fill = Interaction_Difference > 0)) + # Moved inside aes()
  geom_bar(
    stat = "identity", 
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255), 
      "FALSE" = rgb(33/255, 102/255, 172/255)),
    labels = c(
      "TRUE" = "Tumor",
      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(title = "Difference in interactions per source cell (Tumor - Skin)", 
       y = "Difference in Interactions", 
       x = "Source Cell") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

difference_plot

ggsave("~/rocker/RDEB/thesis/count_source_absolute_difference_tumor_vs_skin.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```


Scatter plot source
```{r}
scatter_plot <- ggplot(
  combined_data.diff, 
  aes(x = Tumor_Interactions, 
      y = Skin_Interactions, 
      label = source)) +
  geom_point(
    aes(color = source), size = 3
    ) +
  geom_text(
    aes(label = source), size = 3
    ) +
  geom_smooth(
    method = "lm",
    se = TRUE
    ) +
  labs(
    title = "Number of Interactions per source Cell", 
    x = "Tumor Interactions", 
    y = "Skin Interactions"
    ) +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)

ggsave("~/rocker/RDEB/thesis/scatter_plot_source.png", 
       plot = scatter_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Number of interactions per target cell (receptor) in tumor and skin
```{r}
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>%
  dplyr::count(target) %>%
  dplyr::rename(Number_of_Interactions = n)

df.t$Type <- "Tumor"

df.s <- cellchat.s %>%
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Number_of_Interactions = n)

# Making counts negative
df.s$Number_of_Interactions <- -df.s$Number_of_Interactions 
df.s$Type <- "Skin"

combined_data <- bind_rows(df.t, df.s)

# Plot tumor + skin
combined_plot <- ggplot(
  combined_data, 
  aes(
    x = target, 
    y = Number_of_Interactions, 
    fill = Type)
  ) +
  geom_bar(
    stat = "identity", 
    position = "stack"
    ) +
  labs(
    title = "Number of interactions per target cell", 
    y = "Number of Interactions", 
    x = "Target Cell"
    ) +
  scale_fill_manual(
    values = c(
      "Skin" = rgb(33/255, 102/255, 172/255), 
      "Tumor" = rgb(178/255, 24/255, 43/255)
      )
    ) +
    theme_minimal()  +
  theme(axis.text.x = element_text(
    angle = 90, 
    hjust = 1) # hjust = 1 to "press" names to plot
  )
print(combined_plot)

ggsave("~/rocker/RDEB/thesis/count_target_tumor_vs_skin_without_naive_CD4.png", 
       plot = combined_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)

```

Number of interactions per target cell (receptor) in tumor - skin
```{r}
# Compute interaction counts for each dataset
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Tumor_Interactions = n)
df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::count(target) %>% 
  dplyr::rename(Skin_Interactions = n)

# Merge datasets by the 'target' (27 levels - celltypes) column
combined_data.diff <- full_join(
  df.t,
  df.s, 
  by = "target")

# Remove NA values because T naive CD4 in Skin_Interactions = NA
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each target: tumor - skin
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions
print (combined_data.diff, n = 27) # Display the table (screenshot saved)
# Plot tumor - skin with colors for positive and negative differences
difference_plot <- ggplot(
  combined_data.diff, 
  aes(
    x = target, 
    y = Interaction_Difference, 
    fill = Interaction_Difference > 0)
  ) + # Moved inside aes()
  geom_bar(
    stat = "identity", 
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255), 
      "FALSE" = rgb(33/255, 102/255, 172/255)
      ),
                    labels = c(
                      "TRUE" = "Tumor",
                      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(title = "Difference in interactions per target cell (Tumor - Skin)", 
       y = "Difference in Interactions", 
       x = "Target Cell") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

difference_plot

ggsave("~/rocker/RDEB/thesis/count_target_absolute_difference_tumor_vs_skin.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Scatter plot target
```{r}
scatter_plot <- ggplot(combined_data.diff, 
                       aes(
                         x = Tumor_Interactions, 
                         y = Skin_Interactions, 
                         label = target)
                       ) + 
  geom_point(aes(color = target), size = 3) +
  geom_text(aes(label = target), size = 3) +
  geom_smooth(method = "lm",se = TRUE) +
  labs(
    title = "Number of Interactions per target Cell", 
       x = "Tumor Interactions", 
       y = "Skin Interactions"
    ) +
  theme(legend.position = "none")+
    theme_minimal() 

print(scatter_plot)
ggsave("~/rocker/RDEB/thesis/scatter_plot_target.png", 
       plot = scatter_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Plotting top 10 interactions with biggest differences in tumor - skin
```{r}
# Combining data tumor and skin 
df.t <- cellchat.t %>%
  subsetCommunication(thresh = NA) %>% 
  filter(source != naive_CD4_identifier & target != naive_CD4_identifier) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>%
  dplyr::summarise(Tumor_Interactions = n())

df.s <- cellchat.s %>% 
  subsetCommunication(thresh = NA) %>% 
  dplyr::as_tibble() %>% 
  dplyr::group_by(source, target) %>% 
  dplyr::summarise(Skin_Interactions = n())

# Merge datasets by the 'source' and 'target'
combined_data.diff <- full_join(
  df.t, 
  df.s, 
  by = c("source", "target")
  )

# Replace NAs with 0
combined_data.diff$Tumor_Interactions[
  is.na(combined_data.diff$Tumor_Interactions)
  ] <- 0
combined_data.diff$Skin_Interactions[
  is.na(combined_data.diff$Skin_Interactions)
  ] <- 0

# Compute the difference for each source-target pair: skin - tumor
combined_data.diff$Interaction_Difference <- 
  combined_data.diff$Tumor_Interactions  - combined_data.diff$Skin_Interactions

# Filter to get the top 10 interactions by absolute difference
top_10_diff <- combined_data.diff %>%
  dplyr::ungroup() %>% # Calculate differences without grouping by 'source'
  dplyr::arrange(-abs(Interaction_Difference)) %>%
  dplyr::slice_head(n = 10)

print(top_10_diff)

# Plot top 10 interactions with biggest differences 
difference_plot <- ggplot(
  top_10_diff, 
  aes(
    x = paste(source, target, sep = "-"), 
    y = Interaction_Difference,
    fill = Interaction_Difference > 0)
  ) +
  geom_bar(
    stat = "identity",
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255),
      "FALSE" = rgb(33/255, 102/255, 172/255)
      ),
                    labels = c(
                      "TRUE" = "Tumor",
                      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(
    title = "Top 10 Differences in Interactions (Tumor - Skin)", 
    y = "Difference in Interactions", 
    x = "Interaction (Source-target)") +
  theme(axis.text.x = element_text(
    angle = 90, 
    hjust = 1)
    )

print(difference_plot)

ggsave("~/rocker/RDEB/thesis/Top_10_Differences_in_Interactions.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Function to process interactions between sorce and target celltypes 
```{r}
plot_interaction <- function(
    df, 
    source_value, 
    target_value) {
  df_filtered <- df %>%
    filter(source == source_value & target == target_value) %>%
    slice_max(
      order_by = abs(prob.delta), 
      n = 40) %>%
    select(-c(pval.t, pval.s))

  # Create a label from the source and target values
  interaction_label <- paste(
    source_value, 
    target_value, 
    sep = " to ")

  # Pivot and plot using the new label
  df_filtered %>%
    pivot_longer(cols = c("prob.s", "prob.t"), 
                 names_to = "variable", 
                 values_to = "value") %>%
    ggplot(aes(
      x = interaction_name_2, 
      y = value, 
      fill = variable)
      ) +
    geom_col(position = "dodge") +
    scale_fill_manual(
      values = c(
      "prob.s" = rgb(33/255, 102/255, 172/255), 
      "prob.t" = rgb(178/255, 24/255, 43/255)
      )
      ) +
    labs(
      x = interaction_label, 
      y = "Probability Delta", 
      fill = "Probability"
      ) +
    theme(axis.text.x = element_text(
      angle = 90, 
      hjust = 1, 
      vjust = 0.5)
      )
}
```

Plotting top 10 interactions based on abs(interactions difference)
```{r}
# Define the pairs of interactions
interaction_pairs <- data.frame(
  source = c("Fibroblast_32", "Fibroblast_36", "Fibroblast_32", 
             "Endothel_vasc._31", "Fibroblast_32", "Fibroblast_20", 
             "Fibroblast_32", "Fibroblast_18", "Fibroblast_18", "Endothel"),
  target = c("Fibroblast_32", "Fibroblast_36", "Endothel", "Fibroblast_32",
             "Monocyte", "Endothel", "Fibroblast_20", "Endothel",
             "Fibroblast_36", "Endothel")
)

# Loop over the defined interaction pairs
for (i in 1:nrow(interaction_pairs)) {
  current_source <- interaction_pairs$source[i]
  current_target <- interaction_pairs$target[i]

  p <- plot_interaction(
    merged_data, 
    current_source, 
    current_target
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(
        angle = 90, 
        vjust = 0.5, 
        hjust = 1)
      )
  
    print(p)

# Save the plot
  file_name <- paste0("interaction",
                      "_", 
                      current_source, 
                      "_", 
                      current_target, 
                      ".pdf")
ggsave(file_name,
       plot = p,
       width = 18,
       height = 10,
       dpi = 300,
       path = "~/rocker/RDEB/thesis/")
}
```

Sorting interactions in tumor based on pval and prob
```{r}
# Identifying the top interactions based on the p-value (interactions with pval == 0)
# Group by interaction_name, then filter where pval in the group are 0
interactions_pval_zero.t <- cellchat.t.sC %>%
  group_by(interaction_name) %>%
  filter(pval == 0)

# Calculate the number of interactions to select for the top 1%
top_1_percent_count <- ceiling(nrow(interactions_pval_zero.t) * 0.01) 
# Use ceiling() to round up

# Create a new dataframe sorted by prob and select top 1%
interactions_sorted_prob.t <- interactions_pval_zero.t %>%
  ungroup() %>% # Ungroup to ensure slice_max works as expected
  slice_max(
    order_by = prob,
    n = top_1_percent_count
    )
  
# all interactions summarised, so each interaction = row
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
Top_1_Interactions_in_Tumor <- ggplot(
  interactions_sorted_prob.t, 
  aes(
    x = source, 
    y = target, 
    label = interaction_name)
  ) +
  geom_tile(
    aes(fill = interaction_name), 
    colour = "white"
    ) +
  labs(
    title = "Top 1% Interactions in Tumor", 
    x = "Source", 
    y = "Target", 
    fill = "Interactions"
    ) +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
      )
  )

print(Top_1_Interactions_in_Tumor)

ggsave("~/rocker/RDEB/thesis/Top_1%_Interactions_in_Tumor.png",
       plot = Top_1_Interactions_in_Tumor, 
       width = 10, 
       height = 8,
       dpi = 300)

# Retrieve the full data for these top interactions
details_top_interactions_t <- cellchat.t.sC %>%
  filter(interaction_name %in% interactions_sorted_prob.t$interaction_name)
#  interactions are NOT summarised, so each interaction = multiple rows
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
Top_1_Interactions_in_Tumor_full <- ggplot(
  details_top_interactions_t, 
  aes(
    x = source, 
    y = target, 
    label = interaction_name)
  ) +
  geom_tile(
    aes(fill = interaction_name),
    colour = "white"
    ) +
  labs(
    title = "Top 1% Interactions in Tumor", 
    x = "Source", 
    y = "Target", 
    fill = "Interactions") +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

print(Top_1_Interactions_in_Tumor_full)

ggsave("~/rocker/RDEB/thesis/Top_1%_Interactions_in_Tumor_full.png", 
       plot = Top_1_Interactions_in_Tumor_full, 
       width = 10,
       height = 8,
       dpi = 300)
```

Sorting interactions in skin based on pval and prob
```{r}
# Identifying the top interactions based on the p-value (interactions with pval == 0)
# Group by interaction_name, then filter where pval in the group are 0
interactions_pval_zero.s <- cellchat.s.sC %>%
  group_by(interaction_name) %>%
  filter(pval == 0)
  
# Calculate the number of interactions to select for the top 1%
top_1_percent_count <- ceiling(
  nrow(interactions_pval_zero.s) * 0.01
  ) # Use ceiling to round up
# Create a new dataframe sorted by prob and select top 1%
interactions_sorted_prob.s <- interactions_pval_zero.s %>%
  ungroup() %>% # Ungroup to ensure slice_max works as expected
  slice_max(
    order_by = prob, 
    n = top_1_percent_count
    )
  
# all interactions summarised, so each interaction = row
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
Top_1_Interactions_in_Skin <- ggplot(
  interactions_sorted_prob.s, 
  aes(
    x = source, 
    y = target, 
    label = interaction_name)
  ) +
  geom_tile(
    aes(fill = interaction_name), 
    colour = "white") +
  labs(
    title = "Top 1% Interactions in Skin", 
    x = "Source", 
    y = "Target", 
    fill = "Interactions"
    ) +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1)
  )

print(Top_1_Interactions_in_Skin)

ggsave("~/rocker/RDEB/thesis/Top_1%_Interactions_in_Skin.png", 
       plot = Top_1_Interactions_in_Skin, 
       width = 10, 
       height = 8, 
       dpi = 300)
  
# Retrieve the full data for these top interactions
details_top_interactions_s <- cellchat.s.sC %>%
  filter(interaction_name %in% interactions_sorted_prob.s$interaction_name)
#  interactions are NOT summarised, so each interaction = multiple rows
# 'source' and 'target' are discrete factors, and 'ligand' and 'receptor' are characters
Top_1_Interactions_in_Skin_full <- ggplot(
  details_top_interactions_s, 
  aes(x = source, 
      y = target, 
      label = interaction_name)
  ) +
  geom_tile(
    aes(fill = interaction_name),
    colour = "white"
    ) +
  labs(
    title = "Top 1% Interactions in Skin", 
    x = "Source", 
    y = "Target", 
    fill = "Interactions"
    ) +
  theme_minimal() + #white background 
  theme(
    axis.text.x = element_text(
      angle = 90, 
      hjust = 1
      )
  )

print(Top_1_Interactions_in_Skin_full)

ggsave("~/rocker/RDEB/thesis/Top_1%_Interactions_in_Skin_full.png", 
       plot = Top_1_Interactions_in_Skin_full, 
       width = 10,
       height = 8, 
       dpi = 300)
```

Top 100 Interactions with Biggest Difference in Probability Tumor
```{r}
# Filtering by pval == 0 and abs(prob.delta)
merged_common_data <- merged_data %>%
  filter(pval.t == 0 & pval.s == 0) %>%
  mutate(
    more_common_in = ifelse(
      prob.delta > 0,
      "Tumor",
      "Skin"
      )
  )%>%
  slice_max(
    order_by = abs(prob.delta),
    n = 100)

# Plotting
Top_100_Interactions_with_Biggest_Difference_in_Probability <- ggplot(
  merged_common_data, 
  aes(
    x = source, 
    y = target)
  ) +
  geom_tile(
    aes(fill = more_common_in), 
    colour = "white"
    ) +
  geom_text(
    aes(label = interaction_name),
    size = 3, 
    check_overlap = TRUE,
    vjust = "top"
    ) + # Add interaction names
  scale_fill_manual(
    values = c(
    "Tumor" = rgb(178/255, 24/255, 43/255), 
    "Skin" = rgb(33/255, 102/255, 172/255)
    )
    ) + # Define custom colors for tumor and skin
  labs(
    title = "Top 100 Interactions with Biggest Difference in Probability (Tumor vs. Skin)", 
    x = "Source", 
    y = "Target", 
    fill = "More Common In") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(
      angle = 90, 
      hjust = 1
      ),
    strip.background = element_blank(),
    strip.text = element_text(
      size = 12, 
      face = "bold"
      )
  )
print(Top_100_Interactions_with_Biggest_Difference_in_Probability)
ggsave("~/rocker/RDEB/thesis/Top_100_Interactions_with_Biggest_Difference_in_Probability.png",
       plot = Top_100_Interactions_with_Biggest_Difference_in_Probability, 
       width = 10, 
       height = 8, 
       dpi = 300)
```

Interaction based on cell types
```{r}
cellchat.t@netP$pathways
pathways.show.t <- c("COLLAGEN", "LAMININ", "MIF", "MHC-I", "FN1", "CD99", "THBS", "VISFATIN", "MHC-II", "APP", "CLEC", "MK", "ADGRE5", "GALECTIN", "CXCL", "TGFb", "PARs", "SEMA4", "PTPRM", "SEMA3", "TENASCIN", "CD45", "ANNEXIN", "PTN", "ICAM", "ITGB2", "GAS", "EGF", "PECAM1", "ANGPTL", "TNF", "CD22", "JAM", "CD46", "VEGF", "NOTCH", "THY1", "PERIOSTIN", "CCL", "SELE", "MPZ", "ncWNT", "NEGR", "ALCAM", "CD6", "ANGPT", "NECTIN", "COMPLEMENT", "VCAM", "SELL", "FGF", "PDGF", "PROS", "CADM", "HSPG", "CALCR", "CD226", "BAFF", "CD86", "NCAM", "IL16", "IL1", "SEMA6", "EPHB", "BMP", "EPHA", "LCK", "TIGIT", "IFN-II", "GRN", "ESAM", "NRXN", "KIT", "CDH5", "CSF", "SELPLG", "CDH", "EDN", "BTLA", "IL2", "CNTN", "CD70", "TWEAK", "IGF", "APRIL", "TRAIL", "VISTA", "NRG", "APJ", "RELN", "IL6", "CD34", "HGF", "L1CAM", "SN", "CD137", "WNT")

cellchat.s@netP$pathways
pathways.show.s <- c("COLLAGEN", "LAMININ", "MHC-I", "MIF", "FN1", "CD99", "THBS", "VISFATIN", "CLEC", "APP", "ADGRE5", "MHC-II", "CXCL", "ICAM", "SELE", "MK", "CD45", "PTN", "PTPRM", "ITGB2", "PARs", "CD22", "ALCAM", "CD6", "TGFb", "ANGPTL", "VEGF", "JAM", "CD46", "ANNEXIN", "PECAM1", "GAS", "FGF", "SEMA3", "MPZ", "CADM", "NOTCH", "TENASCIN", "VCAM", "NEGR", "PERIOSTIN", "NECTIN", "IL6", "ANGPT", "SEMA4", "THY1", "EGF", "TIGIT", "TNF", "PDGF", "CCL", "SEMA7", "CALCR", "NCAM", "IL2", "ncWNT", "IFN-II", "EPHA", "PROS", "IL16", "LCK", "EPHB", "CD226", "HSPG", "SELL", "KIT", "BMP", "ESAM", "CD70", "CDH5", "LIGHT", "CD86", "DESMOSOME", "BAFF", "SEMA6", "TRAIL", "COMPLEMENT", "IGF", "CDH1", "SELPLG", "IL1", "FASLG", "EDN", "TWEAK", "NRXN", "IL10", "CD34", "CSF3", "LIFR", "CSF", "NRG", "L1CAM", "CDH", "WNT", "VEGI")

# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, 
           target == current_target)

  # Plot
  p <- ggplot(
    current_data, 
    aes(
      x = ligand, 
      y = receptor, 
      fill = prob.delta)
    ) +
    geom_tile() +
    scale_fill_gradient2(
      low = rgb(33/255, 102/255, 172/255), 
      high = rgb(178/255, 24/255, 43/255), 
      mid = "grey", 
      midpoint = 0,
      limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                 max(current_data$prob.delta, na.rm = TRUE)),
      name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for",
                       current_source,
                       "to",
                       current_target),
         x = "Ligand",
         y = "Receptor"
         ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
      angle = 90, 
      vjust = 0.5
      ),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("~/rocker/RDEB/thesis/interaction_plots/",
                      current_source,
                      "_to_",
                      current_target, 
                      ".pdf"
                      )
  ggsave(
    file_name,
    plot = p,
    width = 10,
    height = 10
    )
  
  # Extracting the relevant interactions from merged_data for the current interaction pair
  filtered_data <- current_data %>%
    distinct(pathway_name)  # Getting unique pathway names involved in these interactions

  # Print the pathways involved in the current interaction pair
  cat("Pathways for interaction", current_source, "to", current_target, ":\n")
  print(filtered_data)

  # Check if these pathways are in the provided pathways lists
  pathways_involved_t <- filtered_data$pathway_name %in% pathways.show.t
  pathways_involved_s <- filtered_data$pathway_name %in% pathways.show.s

  # Print pathways involved both in tumor and skin for the current interaction pair
  cat("Pathways in Tumor:\n")
  print(filtered_data$pathway_name[pathways_involved_t])
  cat("Pathways in Skin:\n")
  print(filtered_data$pathway_name[pathways_involved_s])
}

# List of interactions
interactions <- list(
  "Fibroblast_32 to Fibroblast_32" = list(
    "tumor" = c("COLLAGEN", "NEGR", "MK", "LAMININ", "TENASCIN", "SEMA4", "FN1", "VISFATIN", "FGF", "THBS", "ncWNT", "ANGPTL", "TGFb", "GAS", "CD99", "PTPRM", "JAM", "CXCL", "PERIOSTIN", "SEMA3", "PROS", "CADM", "PDGF", "CNTN", "PTN", "NECTIN", "TWEAK", "BMP", "MIF", "MPZ", "IGF", "WNT"),
    "skin" = c("COLLAGEN", "NEGR", "MK", "LAMININ", "TENASCIN", "SEMA4", "FN1", "VISFATIN", "FGF", "THBS", "ncWNT", "ANGPTL", "TGFb", "GAS", "CD99", "PTPRM", "JAM", "CXCL", "PERIOSTIN", "SEMA3", "PROS", "CADM", "PDGF", "PTN", "NECTIN", "TWEAK", "BMP", "MIF", "MPZ", "IGF", "WNT")
  ),
  "Fibroblast_36 to Fibroblast_36" = list(
    "tumor" = c("COLLAGEN", "ANGPTL", "LAMININ", "MK", "THBS", "MPZ", "VISFATIN", "JAM", "NEGR", "IL6", "SEMA3", "FN1", "SEMA4", "FGF", "GAS", "TENASCIN", "ADGRE5", "PROS", "CXCL", "NOTCH", "CD46", "NECTIN", "PTN", "TWEAK", "CD99", "PERIOSTIN", "HSPG", "PTPRM", "EPHA", "TGFb", "PDGF", "EPHB", "CADM", "MIF", "IGF", "WNT"),
    "skin" = c("COLLAGEN", "ANGPTL", "LAMININ", "MK", "THBS", "MPZ", "VISFATIN", "JAM", "NEGR", "IL6", "SEMA3", "FN1", "SEMA4", "FGF", "GAS", "TENASCIN", "ADGRE5", "PROS", "CXCL", "NOTCH", "CD46", "NECTIN", "PTN", "TWEAK", "CD99", "PERIOSTIN", "HSPG", "PTPRM", "EPHA", "TGFb", "PDGF", "EPHB", "CADM", "MIF", "IGF", "WNT")
  ),
  "Fibroblast_32 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "MK", "VISFATIN", "MIF", "FN1", "CXCL", "SEMA4", "CALCR", "ANGPTL", "CD46", "CCL", "PTN", "ncWNT", "VEGF", "PTPRM", "CD99", "SEMA3", "NECTIN", "APJ", "PDGF", "APP", "IGF"),
    "skin" = c("COLLAGEN", "LAMININ", "MK", "VISFATIN", "MIF", "FN1", "CXCL", "SEMA4", "CALCR", "ANGPTL", "CD46", "CCL", "PTN", "ncWNT", "VEGF", "PTPRM", "CD99", "SEMA3", "NECTIN", "PDGF", "APP", "IGF")
  ),
  "Endothel_vasc._31 to Fibroblast_32" = list(
    "tumor" = c("COLLAGEN", "SELE", "LAMININ", "VISFATIN", "JAM", "ANGPT", "THBS", "GALECTIN", "SEMA4", "CD99", "SEMA6", "NOTCH", "PTPRM", "FN1", "EDN", "TENASCIN", "TGFb", "PROS", "PERIOSTIN", "SEMA3", "VCAM", "CXCL", "BMP", "MIF", "GAS", "MPZ", "EGF", "PDGF", "WNT"),
    "skin" = c("COLLAGEN", "SELE", "LAMININ", "VISFATIN", "JAM", "ANGPT", "THBS", "SEMA4", "CD99", "SEMA6", "NOTCH", "PTPRM", "FN1", "EDN", "TENASCIN", "TGFb", "PROS", "PERIOSTIN", "SEMA3", "VCAM", "CXCL", "BMP", "MIF", "GAS", "MPZ", "EGF", "PDGF", "WNT") 
  ),
  "Fibroblast_32 to Monocyte" = list(
    "tumor" = c("FN1", "COLLAGEN", "CD99", "LAMININ", "COMPLEMENT", "TENASCIN", "MK", "THY1", "MIF", "THBS", "CXCL", "ANNEXIN", "IL6", "FGF", "ANGPTL", "GAS", "CSF", "MHC-I", "SEMA3", "SEMA4", "MHC-II", "JAM", "MPZ", "IL16", "ICAM", "PROS", "NECTIN", "PTN", "APP", "TGFb", "VISFATIN", "VEGF"),
    "skin" = c("FN1", "COLLAGEN", "CD99", "LAMININ", "COMPLEMENT", "TENASCIN", "MK", "THY1", "MIF", "THBS", "CXCL", "ANNEXIN", "IL6", "FGF", "ANGPTL", "GAS", "CSF", "MHC-I", "SEMA3", "SEMA4", "MHC-II", "JAM", "MPZ", "IL16", "ICAM", "PROS", "NECTIN", "PTN", "APP", "TGFb", "VISFATIN", "VEGF")
  ),
  "Fibroblast_20 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "CD99", "VISFATIN", "MIF", "FN1", "VEGF", "CD46", "SEMA4", "ADGRE5", "CCL", "ESAM", "PDGF", "ANGPT", "APP", "PTPRM"),
    "skin" = c("COLLAGEN", "LAMININ", "CD99", "VISFATIN", "MIF", "FN1", "VEGF", "CD46", "SEMA4", "ADGRE5", "CCL", "ESAM", "PDGF", "ANGPT", "APP", "PTPRM")
  ),
  "Fibroblast_32 to Fibroblast_20" = list(
    "tumor" = c("LAMININ", "MK", "CD99", "CD46", "COLLAGEN", "GAS", "PERIOSTIN", "THBS", "VISFATIN", "ncWNT", "NEGR", "CALCR", "MPZ", "HSPG", "JAM", "FN1", "PTN", "PROS", "NECTIN", "PDGF", "CSF", "PTPRM", "BMP", "EPHA", "IGF"),
    "skin" = c("LAMININ", "MK", "CD99", "CD46", "COLLAGEN", "GAS", "PERIOSTIN", "THBS", "VISFATIN", "ncWNT", "NEGR", "CALCR", "MPZ", "HSPG", "JAM", "FN1", "PTN", "PROS", "NECTIN", "PDGF", "CSF", "PTPRM", "BMP", "EPHA", "IGF")
  ),
  "Fibroblast_18 to Endothel" = list(
    "tumor" = c("COLLAGEN", "LAMININ", "VISFATIN", "FN1", "ncWNT", "SEMA4", "CXCL", "VEGF", "ANGPTL", "MIF", "APP", "CD46", "SEMA3", "PTN", "MK", "CCL", "NECTIN", "PTPRM", "APJ", "CALCR", "PDGF", "CD99", "IGF"),
    "skin" = c("COLLAGEN", "LAMININ", "VISFATIN", "FN1", "ncWNT", "SEMA4", "CXCL", "VEGF", "ANGPTL", "MIF", "APP", "CD46", "SEMA3", "PTN", "MK", "CCL", "NECTIN", "PTPRM", "CALCR", "PDGF", "CD99", "IGF")
  ),
  "Fibroblast_18 to Fibroblast_36" = list(
    "tumor" = c("COLLAGEN", "THBS", "LAMININ", "ANGPTL", "FGF", "ncWNT", "VISFATIN", "SEMA4", "IL6", "NEGR", "SEMA3", "GAS", "MPZ", "FN1", "CADM", "TENASCIN", "PTN", "TGFb", "JAM", "MK", "PROS", "PTPRM", "EPHA", "NECTIN", "CD46", "PDGF", "HSPG", "CD99", "MIF", "PERIOSTIN", "CXCL", "IGF"),
    "skin" = c("COLLAGEN", "THBS", "LAMININ", "ANGPTL", "FGF", "ncWNT", "VISFATIN", "SEMA4", "IL6", "NEGR", "SEMA3", "GAS", "MPZ", "FN1", "CADM", "TENASCIN", "PTN", "TGFb", "JAM", "MK", "PROS", "PTPRM", "EPHA", "NECTIN", "CD46", "PDGF", "HSPG", "CD99", "MIF", "PERIOSTIN", "CXCL", "IGF")
  ),
  "Endothel to Endothel" = list(
    "tumor" = c("COLLAGEN", "MIF", "LAMININ", "VISFATIN", "CD99", "SEMA4", "FN1", "PECAM1", "CXCL", "CD46", "CCL", "APP", "ESAM", "PTPRM", "ANGPT", "CDH5"),
    "skin" = c("COLLAGEN", "MIF", "LAMININ", "VISFATIN", "CD99", "SEMA4", "FN1", "PECAM1", "CXCL", "CD46", "CCL", "APP", "ESAM", "PTPRM", "ANGPT", "CDH5")
  )
)

# Find common pathways for each interaction pair
common_pathways_each_pair <- lapply(interactions, function(x) {
  intersect(x$tumor, x$skin)
})

# Find common pathways across all interaction pairs
common_pathways_across_all <- Reduce(intersect, common_pathways_each_pair)

# Print common pathways across all interaction pairs
cat("Common pathways across all interaction pairs:\n")
print(common_pathways_across_all)

# Find pathways that are present in either tumor OR skin for each interaction pair
all_pathways_each_pair <- lapply(interactions, function(x) {
  union(x$tumor, x$skin)
})

# Print pathways for each interaction pair
for (interaction in names(all_pathways_each_pair)) {
  cat("Pathways for interaction", interaction, ":\n")
  print(all_pathways_each_pair[[interaction]])
  cat("\n")
}

# Find unique pathways across all interaction pairs
all_unique_pathways <- unique(unlist(all_pathways_each_pair))

# Print all unique pathways across all interaction pairs
cat("All unique pathways across all interaction pairs:\n")
print(all_unique_pathways)

```

Violin plots
```{r}
# Directory to save the results
output_dir <- "~/rocker/RDEB/thesis/violin_plots/"

# Function to plot and save gene expression for a given pathway
plot_and_save <- function(cellchat_object, signaling, tissue, output_dir) {
  plot_obj <- plotGeneExpression(cellchat_object, signaling = signaling)
  plot(plot_obj)
  file_name <- file.path(output_dir, paste0(signaling, "_", tissue, ".png"))
  ggsave(file_name, plot = plot_obj, width = 10, height = 10)
}

# Plot and save gene expression for each common pathway in tumor and skin
for (pathway in all_unique_pathways) {
  plot_and_save(cellchat.t, signaling = pathway, tissue = "tumor", output_dir)
  plot_and_save(cellchat.s, signaling = pathway, tissue = "skin", output_dir)
}
```

























bei logFC auch auf difference aufpassen, so that not only statistics look good, but also correlates withe biology

1) Circle diagramm for skin, tumor
```{r}
# Calculate the aggregated cell-cell communication network tumor
# Open PDF device
pdf("~/rocker/RDEB/thesis/NetworkVisualizations_allTumor.pdf", width = 11, height = 8)

# Set up plotting parameters
groupSize <- as.numeric(table(cellchat.t@idents))
par(mfrow = c(1, 2), xpd = TRUE, mar = c(4, 4, 2, 1))

# Generate the plots
netVisual_circle(cellchat.t@net$count, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = "Number of interactions")
netVisual_circle(cellchat.t@net$weight, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = "Interaction weights/strength")

# Due to the complicated cell-cell communication network, we can examine the signaling sent from each cell group. 
# Here we also control the parameter edge.weight.max so that we can compare edge weights between differet networks.
# Saving plots to a PDF
pdf("~/rocker/RDEB/thesis/NetworkVisualizations_Tumor.pdf", width = 14, height = 10)

par(mfrow = c(3, 4), mar = c(2, 2, 2, 1), xpd = TRUE)
mat <- cellchat.t@net$weight
groupSize <- rep(10, nrow(mat))

for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = TRUE,
                   edge.weight.max = max(mat, na.rm = TRUE), title.name = rownames(mat)[i])
}

# Output PDF setup
pdf("~/rocker/RDEB/thesis/Chord_diagram_Tumor.pdf")
netVisual_aggregate(cellchat.t, signaling = pathways.show.t, layout = "chord")

# Calculate the aggregated cell-cell communication network skin
# Open PDF device
pdf("~/rocker/RDEB/thesis/NetworkVisualizations_allSkin.pdf", width = 11, height = 8)

# Set up plotting parameters
groupSize <- as.numeric(table(cellchat.s@idents))
par(mfrow = c(1, 2), xpd = TRUE, mar = c(4, 4, 2, 1))  # Adjust margins if necessary

# Generate the plots
netVisual_circle(cellchat.s@net$count, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = "Number of interactions")
netVisual_circle(cellchat.s@net$weight, vertex.weight = groupSize, weight.scale = TRUE, label.edge = FALSE, title.name = "Interaction weights/strength")

# Due to the complicated cell-cell communication network, we can examine the signaling sent from each cell group. 
# Here we also control the parameter edge.weight.max so that we can compare edge weights between differet networks.
# Saving plots to a PDF
pdf("~/rocker/RDEB/thesis/NetworkVisualizations_Skin.pdf", width = 14, height = 10)

par(mfrow = c(3, 4), mar = c(2, 2, 2, 1), xpd = TRUE)
mat <- cellchat.s@net$weight
groupSize <- rep(10, nrow(mat))

for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = TRUE,
                   edge.weight.max = max(mat, na.rm = TRUE), title.name = rownames(mat)[i])
}

# Output PDF setup
pdf("~/rocker/RDEB/thesis/Chord_diagram_Skin.pdf")
netVisual_aggregate(cellchat.s, signaling = pathways.show.s, layout = "chord")
```
  
2) Heatmap tumor, heatmap skin
```{r}
# Set up PDF output for Tumor, saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Tumor.pdf", width = 16, height = 8)

# Interaction count heatmap for Tumor
gg1_tumor <- netVisual_heatmap(cellchat.t, measure = "count")

# Interaction weight heatmap for Tumor
gg2_tumor <- netVisual_heatmap(cellchat.t, measure = "weight")

# Combine and plot the heatmaps side by side for Tumor
gg_combined_tumor <- gg1_tumor + gg2_tumor
plot(gg_combined_tumor)

# Set up PDF output for Skin, also saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Skin.pdf", width = 16, height = 8)

# Interaction count heatmap for Skin
gg1_skin <- netVisual_heatmap(cellchat.s, measure = "count")

# Interaction weight heatmap for Skin
gg2_skin <- netVisual_heatmap(cellchat.s, measure = "weight")

# Combine and plot the heatmaps side by side for Skin
gg_combined_skin <- gg1_skin + gg2_skin
plot(gg_combined_skin)


# Set up PDF output for Tumor, saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Tumor_Interaction_count.pdf", width = 16, height = 8)

# Interaction count heatmap for Tumor
gg1_tumor <- netVisual_heatmap(cellchat.t, measure = "count")
plot(gg1_tumor)

# Set up PDF output for Tumor, saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Tumor_Interaction_weight.pdf", width = 16, height = 8)

# Interaction weight heatmap for Tumor
gg2_tumor <- netVisual_heatmap(cellchat.t, measure = "weight")
plot(gg2_tumor)

# Set up PDF output for Skin, also saving directly to the desired directory
pdf("~/rocker/RDEB/thesis/Heatmap_Skin_Interaction_count.pdf", width = 16, height = 8)

# Interaction count heatmap for Skin
gg1_skin <- netVisual_heatmap(cellchat.s, measure = "count")

plot(gg1_skin)

pdf("~/rocker/RDEB/thesis/Heatmap_Skin_Interaction_weight.pdf", width = 16, height = 8)

# Interaction weight heatmap for Skin
gg2_skin <- netVisual_heatmap(cellchat.s, measure = "weight")

plot(gg2_skin)
```

Here i am creating new grafics for abs(Interaction_Difference_abs) top10 interactions
```{r}
merged_df.interactions

# Filter to get the top 10 interactions by absolute difference
top_10_diff <- merged_df.interactions%>%
  dplyr::ungroup() %>% # Calculate differences without grouping by 'source'
  dplyr::arrange(-abs(Interaction_Difference_abs)) %>%
  dplyr::slice_head(n = 10)

print(top_10_diff)

# Plot top 10 interactions with biggest differences 
difference_plot <- ggplot(
  top_10_diff, 
  aes(
    x = paste(source, target, sep = "-"), 
    y = Interaction_Difference_abs,
    fill = Interaction_Difference_abs > 0)
  ) +
  geom_bar(
    stat = "identity",
    position = "dodge"
    ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rgb(178/255, 24/255, 43/255),
      "FALSE" = rgb(33/255, 102/255, 172/255)
      ),
                    labels = c(
                      "TRUE" = "Tumor",
                      "FALSE" = "Skin")
    ) +
  theme_minimal() +
  labs(
    title = "Top 10 according to Interaction_Difference_abs (Tumor - Skin)", 
    y = "Difference in Interactions", 
    x = "Interaction (Source-target)") +
  theme(axis.text.x = element_text(
    angle = 90, 
    hjust = 1)
    )

print(difference_plot)

ggsave("~/rocker/RDEB/thesis/Top_10_Interaction_Difference_abs.png", 
       plot = difference_plot, 
       width = 10, 
       height = 8, 
       dpi = 300)

# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, 
           target == current_target)

  # Plot
  p <- ggplot(
    current_data, 
    aes(
      x = ligand, 
      y = receptor, 
      fill = prob.delta)
    ) +
    geom_tile() +
    scale_fill_gradient2(
      low = rgb(33/255, 102/255, 172/255), 
      high = rgb(178/255, 24/255, 43/255), 
      mid = "grey", 
      midpoint = 0,
      limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                 max(current_data$prob.delta, na.rm = TRUE)),
      name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for",
                       current_source,
                       "to",
                       current_target),
         x = "Ligand",
         y = "Receptor"
         ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
      angle = 90, 
      vjust = 0.5
      ),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("~/rocker/RDEB/thesis/interaction_plots_abs/",
                      current_source,
                      "_to_",
                      current_target, 
                      "_abs_",
                      ".pdf"
                      )
  ggsave(
    file_name,
    plot = p,
    width = 10,
    height = 10
    )
  
  # Extracting the relevant interactions from merged_data for the current interaction pair
  filtered_data <- current_data %>%
    distinct(pathway_name)  # Getting unique pathway names involved in these interactions

  # Print the pathways involved in the current interaction pair
  cat("Pathways for interaction", current_source, "to", current_target, ":\n")
  print(filtered_data)

  # Check if these pathways are in the provided pathways lists
  pathways_involved_t <- filtered_data$pathway_name %in% pathways.show.t
  pathways_involved_s <- filtered_data$pathway_name %in% pathways.show.s

  # Print pathways involved both in tumor and skin for the current interaction pair
  cat("Pathways in Tumor:\n")
  print(filtered_data$pathway_name[pathways_involved_t])
  cat("Pathways in Skin:\n")
  print(filtered_data$pathway_name[pathways_involved_s])
}


# Filter to get the top 10 interactions by absolute difference, considering both parameters
top_10_diff <- merged_df.interactions %>%
  dplyr::ungroup() %>%
  dplyr::arrange(desc(abs(Interaction_Difference)), desc(abs(Interaction_Difference_abs))) %>%
  dplyr::slice_head(n = 20)

print(top_10_diff)

# Now, let's loop over these top 10 interactions and create the desired plots.
for (i in 1:nrow(top_10_diff)) {
  current_source <- top_10_diff$source[i]
  current_target <- top_10_diff$target[i]

  # Filter the merged_data for the current pair.
  current_data <- merged_data %>%
    filter(source == current_source, 
           target == current_target)

  # Plot
  p <- ggplot(
    current_data, 
    aes(
      x = ligand, 
      y = receptor, 
      fill = prob.delta)
    ) +
    geom_tile() +
    scale_fill_gradient2(
      low = rgb(33/255, 102/255, 172/255), 
      high = rgb(178/255, 24/255, 43/255), 
      mid = "grey", 
      midpoint = 0,
      limits = c(min(current_data$prob.delta, na.rm = TRUE), 
                 max(current_data$prob.delta, na.rm = TRUE)),
      name = "Prob Delta") +
    labs(title = paste("Ligand-Receptor Interactions for",
                       current_source,
                       "to",
                       current_target),
         x = "Ligand",
         y = "Receptor"
         ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
      angle = 90, 
      vjust = 0.5
      ),
          axis.text.y = element_text(hjust = 1))
   
  print(p)

  # Save the plot
  file_name <- paste0("~/rocker/RDEB/thesis/interaction_plots_abs_both/",
                      current_source,
                      "_to_",
                      current_target, 
                      "_abs_",
                      ".pdf"
                      )
  ggsave(
    file_name,
    plot = p,
    width = 10,
    height = 10
    )
  
  # Extracting the relevant interactions from merged_data for the current interaction pair
  filtered_data <- current_data %>%
    distinct(pathway_name)  # Getting unique pathway names involved in these interactions

  # Print the pathways involved in the current interaction pair
  cat("Pathways for interaction", current_source, "to", current_target, ":\n")
  print(filtered_data)

  # Check if these pathways are in the provided pathways lists
  pathways_involved_t <- filtered_data$pathway_name %in% pathways.show.t
  pathways_involved_s <- filtered_data$pathway_name %in% pathways.show.s

  # Print pathways involved both in tumor and skin for the current interaction pair
  cat("Pathways in Tumor:\n")
  print(filtered_data$pathway_name[pathways_involved_t])
  cat("Pathways in Skin:\n")
  print(filtered_data$pathway_name[pathways_involved_s])
}
```